### 2023-01-08 LICS2023.tex
Add files via upload

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usetikzlibrary{cd}
\usepackage{stmaryrd}

\input{macros}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Tropical Geometry and the Lambda Calculus}

\author{\IEEEauthorblockN{Davide Barbarossa}
\IEEEauthorblockA{\textit{Universit\`a di Bologna}\\
Bologna, Italy \\
davide.barbarossa2@unibo.it}
\and
\IEEEauthorblockN{Paolo Pistone}
\IEEEauthorblockA{\textit{Universit\`a Roma Tre}\\
Rome, Italy\\
paolo.pistone@uniroma3.it}
}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
\end{IEEEkeywords}

\section{Introduction}

In the recent years, more and more interest in the programming language community has been directed towards the analysis of \emph{quantitative} properties of algorithms, such as e.g.~number of computation steps, convergence probability, 
approximation bounds, 
as opposed to purely \emph{qualitative} properties like e.g.~termination or program equivalence.

Notably, a significative effort has been made in order to extend, or adapt, well-established qualitative methods, like type systems, relational logics or denotational semantics, to account for quantitative properties. We can mention, for example, 
intersection type systems aimed at capturing time or space resources \cite{Beniamino}, or convergence probabilities \cite{UgoBreuvart, LICS2022},  relational logics to account for probabilistic properties like e.g.~differential privacy \cite{Barthes} or metric preservation \cite{Reed2010, dallago}, as well as the study of denotational models for 
probabilistic \cite{PCOH, QBS} or differential \cite{difflambda} extensions of the $\lambda$-calculus. 
Usually, the main reason to resort to such logical, type-theoretical or denotational approaches is to obtain methods which are \emph{modular} and \emph{compositional}, that is, where properties of complex algorithms can be deduced from properties of their component parts.

\subsection{Two kinds of quantitative approaches}

Among such quantitative approaches, two different directions have received a considerable attention in recent years.

On the one hand there is the approach based on \emph{non-idempotent} intersection types \cite{} and their semantical counterpart, the so-called \emph{relational semantics} \cite{}, which have been exploited to account for higher-order program differentiation \cite{difflambda}, to establish reasonable \emph{cost-models} for the $\lambda$-calculus \cite{}, and have also been shown suitable to capture probabilistic properties \cite{Manzo}.

On the other hand there is the approach of \emph{program metrics} \cite{} and \emph{quantitative equational theories} \cite{Mardare}: when considering probabilistic or approximated computations, rather than knowing whether two programs compute \emph{the same} function, which is rarely the case, it is often more useful to know whether they compute functions that do not differ \emph{too much}. This has motivated the study of denotational semantics in which types are endowed with a metric, measuring similarity of behavior; this approach has found  applications in e.g.~differential privacy \cite{} and coinductive methods \cite{}, and was recently extended to account for the the full $\lambda$-calculus \cite{LICS2021, FSCD2022}.

In both approaches the notion of \emph{linearity}, in the sense of linear logic \cite{linearlogic} (i.e.~of using its input exactly once), plays a crucial role.
In the relational semantics linear programs correspond to linear maps, in the usual algebraic sense, while the possibility of duplicating inputs leads to consider \emph{analytic} maps, i.e.~formal infinite series of monomials of the form $a_{n}\cdot x^{n}$.
By contrast, in metric semantics linear programs correspond to \emph{non-expansive} maps, i.e.~to functions that do not increase distances, while the possibility of duplicating inputs leads to interpret \emph{bounded} programs (i.e.~programs with a fixed duplication bound) as \emph{Lipschitz-continuous} maps \cite{Gabo2017}, and general programs as maps satisfying coarser 
continuity conditions (see e.g.~\cite{LICS2021}).

A natural question is thus whether these two, apparently unrelated, ways of interpreting linearity and duplication can be somehow connected. At a first glance, there seems to be a fundamental ``logarithmic'' gap between the two approaches.
In the relational semantics a typical ``duplicating'' map is obtained by composing the diagonal with multiplication:
$$
\begin{tikzcd}
\mathbb R \ar{rrr}{x\mapsto \langle x, x\rangle}
 & &  &
 \mathbb R\times \mathbb R 
 \ar{rrr}{\langle x,y\rangle \mapsto x\cdot y}
 & & & \mathbb R
\end{tikzcd}
$$
yielding the square product function $\lambda x.x^{2}$.
However, in metric semantics this function needs not even exist (as these models are often restricted to Lipschitz-continuous maps \cite{Gabo2017})! Instead, a typical ``duplicating'' map can be obtained by composing the diagonal with the sum 
$$
\begin{tikzcd}
\mathbb R \ar{rrr}{x\mapsto \langle x, x\rangle}
 & &  &
 \mathbb R\times \mathbb R 
 \ar{rrr}{\langle x,y\rangle \mapsto x+y}
 & & & \mathbb R
\end{tikzcd}
$$
yielding the linear (and Lipschitz) function $\lambda x.2x$.

As this example seems to suggest, there seems to be a sort of ``logarithmic'' gap between the two approaches. Can this be made explicit?

\subsection{Tropical Geometry: a ``logarithmic'' version of usual geometry}

In very simple words, tropical geometry describes what happens to usual algebra and geometry when we replace ``$\cdot$'' by  ``$+$'', and ``$+$'' by ``$\min$''. It can thus be seen as a sort of ``logarithmic'' version of usual geometry (this idea can be made precise via the so-called \emph{Maslov deformation} \cite{}).
Tropical geometry is a kind of \emph{idempotent} mathematics, since the role of addition is 
played by the idempotent operation $\min$.
For instance, a quadratic polynomial like e.g.~$p(x)=ax^{2}+bx+c$ corresponds in the tropical word correspond to the \emph{linear} function
$$
\trop p(\alpha)=\min\{2\alpha + a, \alpha+b, c\}
$$

The tropicalization of usual algebraic concepts comes with the important advantage that many intractable properties may become efficiently tractable, in a computational way (e.g.~finding roots of polynomials). For this reason, methods from tropical geometry are more and more applied in computer science, e.g.~for convex analysis and machine learning \cite{}.

Coming back to our previous discussion of program semantics, tropical geometry might seem to provide precisely what we are looking for: a way to see polynomials as linear (i.e.~Lipschitz) functions.
At this point, it is worth mentioning that a tropical variant of relational semantics has already been considered \cite{}, and shown capable of capturing \emph{optimal} probabilistic properties, but has not yet been studied in detail. Furthermore, connections between tropical linear algebra and the abstract study of metric spaces and non-expansive functions, and more generally, of the study of \emph{quantale-enriched} categories, have also been observed \cite{}.

The goal of this paper is to make such intuitive connections precise, yielding a way to relate, conceptually, two different quantitative approaches to higher-order algorithms, as well as suggesting new potential applications of tropical geometry to the study of quantitative properties of programs.

\subsection{Contributions}

Our contributions in this paper are threefold:
\begin{itemize}

\item We study the relational model over the tropical semi-ring, showing that
\emph{bounded} simply typed terms correspond to tropical polynomials, while general
 simply typed lambda terms correspond to a generalization of \emph{tropical Laurent series} \cite{}. Moreover, we show that such maps are concave and locally Lipschitz-continuous, yielding a full-scale metric semantics for the $\lambda$-calculus.
Moreover, we exploit the differential structure of the relational model to study the \emph{tropical Taylor expansion} of a $\lambda$-term, which can be seen as an approximation of the term by way of Lipschitz-continuous maps.

\item Using the relational model as our main reference, we suggest a list of potential applications of tropical methods to the study of quantitative properties of non-deterministic/probabilistic functional programs like differential privacy, convergence \emph{log-probabilities} and the number of computation steps in the best case.

\item We conclude by putting the correspondence between differential and metric properties made possible by the tropical viewpoint at the right level of generality. 
By recalling and suitably extending a well-known correspondence between Lawvere's \emph{generalized metric spaces} (GMS) \cite{} and modules over the tropical semi-ring \cite{}, we show that the category of \emph{complete}  GMSs provides a model of the differential $\lambda$-calculus which extends the tropical relational model.
\end{itemize}
%
%\section{Bounded and Differential $\lambda$-Calculi}
%
%
%Bounded Simply Typed $\lambda$-calculus $\BSTLC$:
%$$
%A::= o \mid !_{n}A \multimap A
%$$
%
%
%Resource Simply Typed $\lambda$-calculus $\RSTLC$:
%$$
%A::= o \mid [A, \dots , A] \multimap A
%$$
%
%
%Define a translation of types $(-)^{\C R}$ from $\BSTLC$ to $\RSTLC$ by $o^{\C R}=o$ and $(!_{n}A\multimap B)^{\C R}=
%[\underbrace{A^{\C R},\dots, A^{\C R}}_{n\text{ times}}]\multimap B^{\C R}$.
%
%\begin{proposition}
%$\Gamma \vdash_{\BSTLC} M:A$ implies 
%$\Gamma^{\C R}\vdash_{\RSTLC}M:A^{\C R}$.
%\end{proposition}
%

\section{Tropical Geometry}

- The tropical semi-ring $\Lawv$ coincides with the Lawvere quantale.

- tropical linear algebra: the category $\LREL$

- tropical polynomials and tropical roots

- tropical Laurent series and tropical roots

- tropicalization of analytic functions and their roots: 
$\trop P$

- the differential operator

\section{The Tropical Relational Model}

Describe the model in a parametric way over a continuous semi-ring, as in Manzonetto et al.

Define the tropicalization $\trop$ from the $\BB R$-based model
to the $\Lawv$-based model.

Focus on \emph{bounded} simply typed $\lambda$-calculus $\BSTLC$:
$$
A::= o \mid !_{n}A \multimap A
$$
The interpretation of a bounded type is defined as 
$\model A= \Lawv^{\nodel A}$, where 
$\nodel A$ is defined by $\nodel o=\B 1$ and 
$\nodel{!_{n}A\multimap B}:= \C M_{\leq n}(\nodel A) \times \nodel B$.

Bounded types are interesting
because of the following proposition:
\begin{proposition}
For all bounded types $A,B$, the morphisms from $\model A$ to $\model B$ (in all parametric relational models) are polynomials.

\end{proposition}
\begin{proof}
It suffices to check that $\nodel A$ is finite for all bounded types $A$. Indeed this implies that a morphism $f:\model A\to \model B$ corresponds to a finite matrix $\widehat f: \nodel A \times \nodel B \to \Lawv$.
\end{proof}

For example (here $!_{n}(\Lawv^{X}):= \Lawv^{\C M_{\leq n}(X)}$):
\begin{itemize}
\item a map $f\in \LREL( !_{1}\Lawv, \Lawv)$ is of the form $f(x)=\min \{x+a,b\}$;

\item a map $f\in \LREL(!_{2}\Lawv, \Lawv)$ is a ``quadratic'' polynomial $f(x)=\min\{2x+a, x+b, c\}$.

\end{itemize}

Discuss the tropical Taylor expansion $\trop\C T(M)$.

\section{Analytic Results}

- the metric on the tensor is the usual tensor metric, the metric on the function space is the usual function metric, the metric on multisets is the multiset metric

- concavity

- Lipschitz-continuity. Highlight compositional reasoning based on Lipschitz.

- Discussion of unary tropical Laurent series, i.e.~$f:\Lawv \to \Lawv$:
\begin{itemize}

\item $f$ is locally a polynomial (i.e.~a $\min$) but not globally

\item singular points: the famous example

\item tropical roots

\end{itemize}

- discussion of the tropical Taylor expansion:

- characterization of the functional metric using derivatives

\section{Applications}

General discussion: optimization properties behave in a Lipschitz way.

- differential privacy and Lipschitzness

- log-probabilities and tropical roots

- counting computation steps (from Manzonetto, but add relational ``Lipschitz'' reasoning)

- measuring duplications of discrete functions (needs finiteness!)

\section{Quasi-metric Spaces and $\Lawv$-Modules}

- correspondence between $\Lawv$-modules and $\Lawv$-categories.

- exponential structure of $\LCAT$.

- $\LCAT_{!}$ is a cartesian closed differential category

\section{Related Work}

\section{Conclusion}

\end{document}