### 2023-01-08 section3.tex
section split

Describe the model in a parametric way over a continuous semi-ring, as in Manzonetto et al.

Define the tropicalization $\trop$ from the $\BB R$-based model
to the $\Lawv$-based model.

Focus on \emph{bounded} simply typed $\lambda$-calculus $\BSTLC$:
$$
A::= o \mid !_{n}A \multimap A
$$
The interpretation of a bounded type is defined as 
$\model A= \Lawv^{\nodel A}$, where 
$\nodel A$ is defined by $\nodel o=\B 1$ and 
$\nodel{!_{n}A\multimap B}:= \C M_{\leq n}(\nodel A) \times \nodel B$.

Bounded types are interesting
because of the following proposition:
\begin{proposition}
For all bounded types $A,B$, the morphisms from $\model A$ to $\model B$ (in all parametric relational models) are polynomials.

\end{proposition}
\begin{proof}
It suffices to check that $\nodel A$ is finite for all bounded types $A$. Indeed this implies that a morphism $f:\model A\to \model B$ corresponds to a finite matrix $\widehat f: \nodel A \times \nodel B \to \Lawv$.
\end{proof}

For example (here $!_{n}(\Lawv^{X}):= \Lawv^{\C M_{\leq n}(X)}$):
\begin{itemize}
\item a map $f\in \LREL( !_{1}\Lawv, \Lawv)$ is of the form $f(x)=\min \{x+a,b\}$;

\item a map $f\in \LREL(!_{2}\Lawv, \Lawv)$ is a ``quadratic'' polynomial $f(x)=\min\{2x+a, x+b, c\}$.

\end{itemize}

Discuss the tropical Taylor expansion $\trop\C T(M)$.