### 2023-01-08 section3.tex
commit

- matrices over a continuous semi-ring

- The tropical semi-ring $\Lawv$ coincides with the Lawvere quantale.

- tropical linear algebra: the category $\LREL$

- linear computation inside $\LREL$: linear lambda-terms through the symmetric monoidal adjunction 
$$
\LREL(Z\otimes X, Y) \simeq \LREL (Z, X\multimap Y)
$$
(with $\LREL(X,Y)\simeq \Lawv^{X\times Y}$.