### 2023-01-08 section7.tex

- correspondence between $\Lawv$-modules and $\Lawv$-categories.

- exponential structure of $\LCAT$.

- $\LCAT_{!}$ is a cartesian closed differential category