### 2023-01-09 section3.tex
Nuove modifiche

- $\LREL$ is SMCC (Giulio): linear lambda-terms through the symmetric monoidal adjunction 
$$
\LREL(Z\otimes X, Y) \simeq \LREL (Z, X\multimap Y)
$$
(with $\LREL(X,Y)\simeq \Lawv^{X\times Y}$.