### 2023-01-10 section3A.tex
Push serale
Prime parti di section3.
Inoltre ho aggiunto un \label in main.tex ed qualche macros utile.

[Section III.A, Weighted], immediately gives:

\begin{fact}\label{fact:LREL_SMCC}
 $\LREL$ is a Symmetric Monoidal Closed Category.
\end{fact}

Its monoidal structures is given by a tensor product $\otimes$ acting on the objects as the Cartesian product of sets, and as the Kronecker product of matrices on morphisms.
Its closed structures is given by $X\multimap Y:=X\times Y$ with the usual evaluation and curry morphisms yielding a symmetric monoidal adjunction: $\LREL(Z\otimes X, Y) \simeq \LREL (Z, X\multimap Y)$.

Fact~\ref{fact:LREL_SMCC} means that $\LREL$ is a model of Intuitionistic Multiplicative Linear Logic (IMLL), and thus we can interpret \emph{linear} $\lam$-terms in it.