### 2023-01-12 main.tex
commit

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usetikzlibrary{cd}
\usepackage{mathabx}
\usepackage{cmll}
\usepackage{stmaryrd}

\input{macros}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Tropical Mathematics for the Lambda Calculus}

\author{\IEEEauthorblockN{Davide Barbarossa}
\IEEEauthorblockA{\textit{Universit\`a di Bologna}\\
Bologna, Italy \\
davide.barbarossa@unibo.it}
\and
\IEEEauthorblockN{Paolo Pistone}
\IEEEauthorblockA{\textit{Universit\`a Roma Tre}\\
Rome, Italy\\
paolo.pistone@uniroma3.it}
}

\maketitle

\begin{abstract}

\end{abstract}

\begin{IEEEkeywords}
\end{IEEEkeywords}

\section{Introduction}
\input{section1}

\section{A Bridge Between Two Quantitative Approaches to the $\lambda$-calculus}
\input{section2}

\section{Tropical (non-)Linear Algebra and computation}

\input{section3intro}%

\subsection{As a model of linear programs}
\input{section3A}

\subsection{As a model of non-linear programs}
\input{section3B}%

\subsection{As a model of differential programs}
\input{section3C}%

\section{Tropical Laurent series}
\input{section4}%

\section{Applications}
\input{section5}%

\section{Generalized Metric Spaces and Modules over the Lawvere Quantale}\label{sec:QMod}
\input{section6}%

\section{Related Work}
\input{section7}

\section{Conclusion and future work}
\input{section8}

\end{document}