### 2023-01-19 appV.tex

\subsubsection{Proof of Theorem \ref{theorem:fepsilon}}

Also proof of 
\begin{proposition}\label{prop:nondecr+conc}
 Any tLs $f:\Lawv^X\to\Lawv^Y$ is non-decreasing and concave, w.r.t.\ the pointwise order.
\end{proposition}

\subsubsection{Continuity of tLs}

Continuity of cones and metric

\subsubsection{Lipschitz continuity}

\begin{proposition}\label{prop:troplinear}
All tropical linear functions $f: \Lawv^{X}\to \Lawv^{Y}$ are non-expansive.  
\end{proposition}

\begin{lemma}
Let $f: [0,\infty)^{X}\to Q$ be concave, monotone increasing, and continuous.  
Let $\B x\neq \B y\in [0,\infty)^{X}$, with $\|\B x-\B y\|_{\infty}<\infty$, and let $S(\B x, \B y)=\{ \alpha\B x + (1-\alpha)\B y\mid \alpha \in [0,1]\}$ be the segment generated by $\B x$ and $\B y$. 
Then $f$ is Lipschitz-continuous over $S(\B x, \B y)$.

\end{lemma}
\begin{proof}
Let us prove the lemma under the assumption that for all $a\in X$, $\B y_{a}-\B x_{a}\geq 1$. From the fact that the claim holds under the assumption, we can deduce the claim of the lemma:
indeed for $\alpha\in (0,1)$ large enough we have that $\B y':= \frac{\B y-\alpha\B x}{1-\alpha}$ is such that $\B y\in S(\B x, \B y')$ (and thus $S(\B x, \B y)\subseteq S(\B x, \B y')$) and 
$\B y'_{a}-\B x_{a}\geq 1$. Hence from our proof we deduce that $f$ is Lipschitz-continuous over $S(\B x, \B y')$, and thus a fortiori over $S(\B x, \B y)$ too.

Since $f$ is continuous over $[0,\infty)^{X}$ and $S(\B x, \B y)$ is compact, $f$ admits a maximum $\mathrm{MAX}$ over $S(\B x, \B y)$.
For all $\B z< \B z'\in S(\B x, \B y)$, let $M(\B z, \B z')\in Q^{X}$ be defined by
$$
M(\B z, \B z')_{a}= \frac{f(\B z')-f(\B z)}{\B z'_{a}-\B z_{a}}
$$
Observe that
\begin{align*}
M(\B x, \B y)_{a} & = \frac{f(\B y)-f(\B x)}{\B y_{a}-\B x_{a}} \leq 
f(\B y)-f(\B x) \leq \mathrm{MAX}
\end{align*}
using the fact that $\B y_{a}-\B x_{a}\geq 1$.

We now claim that $M(\B z, \B z')$ is contravariant in both $\B z$ and $\B z'$. 
Indeed suppose $\B z\leq \B z'' < \B z'$, so that $\B z= \lambda \B z' +(1-\lambda)\B z''$ for some $\lambda \in (0,1)$. Then, using the fact that $f$ is concave, we have 
\begin{align*}
M(\B z, \B z')_{a}&=\frac{f(\B z')-f(\lambda \B x +(1-\lambda)\B z'')}{\B z'_{a}-\lambda \B x_{a} -(1-\lambda)\B z''_{a}} \\
&
\geq 
\frac{f(\B z')-\lambda f(\B z') -(1-\lambda)f(\B z'')}{\B z'_{a}-\lambda \B z'_{a} -(1-\lambda)\B z''_{a}} \\
&
= 
\frac{(1-\lambda)(f(\B z')-f(\B z''))}{(1-\lambda )\B z'_{a}-\B z''_{a}} =M(\B z'', \B z')
\end{align*}
In a similar way it is shown that for $\B z < \B z''\leq \B z'$, $M(\B z,\B z')\leq M(\B z, \B z'')$.

Therefore, for all $\B z < \B z'\in S(\B x, \B y)$, we have that $M(\B z, \B z')_{a} \leq M(\B x, \B z')_{a} \leq M(\B x, \B y)_{a} \leq \mathrm{MAX}$. 
From this, using the fact that $f$ is monotone increasing, we deduce that 
$|f(\B z')-f(\B z)|=f(\B z')-f(\B z) \leq \mathrm{MAX}\cdot |\B z'_{a}-\B z_{a}|
$ and thus that 
$$|f(\B z')-f(\B z)|\leq \mathrm{MAX}\cdot \|\B z'-\B z\|_{\infty}$$
that is, that $f$ is $\mathrm{MAX}$-Lipschitz over $S(\B x, \B y)$. 
\end{proof}

\begin{proposition}
Let $f: [0,\infty)^{X}\to Q$ be concave, monotone increasing, and continuous.  
For all $\epsilon \in (0,\infty)$ and $\B x\in [0,\infty)^{X}$, $f$ is Lipschitz-continuous over the open ball $B(\B x, \epsilon)$.
\end{proposition}
\begin{proof}
Let $\mathrm{MAX}$ indicate the maximum of $f$ over $B(\B x, \epsilon)$. 
Let $\B y, \B z\in B(\B x, \epsilon)$; then $\|\B y-\B z\|_{\infty}\leq 2\epsilon < \infty$, so by the lemma above $f$ is $K$-Lipschitz over the segment $S(\B y, \B z)$ for some $K\leq \mathrm{MAX}$, so we deduce $|f(\B y)-f(\B z)|\leq \mathrm{MAX}\cdot \|\B y-\B z\|_{\infty}$.
\end{proof}

\begin{theorem}[local Lipschitz-continuity]
Let $f: [0,\infty)^{X}\to Q$ be concave, monotone increasing, and continuous.  
Then $f$ is locally Lipschitz-continuous.
\end{theorem}
\begin{proof}
For all $\B x\in [0,\infty)^{X}$, $f$ is Lipschitz-continuous over the open set $B(\B x, 1)$.
\end{proof}