### 2023-01-19 main.tex

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{bussproofs}
\EnableBpAbbreviations
\usepackage{hyperref}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
\usepackage{tikz-cd}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usetikzlibrary{cd}
\usepackage{mathabx}
\usepackage{cmll}
\usepackage{stmaryrd}

\input{macros}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Tropical Mathematics for the Lambda Calculus}

%\author{\IEEEauthorblockN{Davide Barbarossa}
%\IEEEauthorblockA{\textit{Universit\`a di Bologna}\\
%Bologna, Italy \\
%davide.barbarossa@unibo.it}
%\and
%\IEEEauthorblockN{Paolo Pistone}
%\IEEEauthorblockA{\textit{Universit\`a Roma Tre}\\
%Rome, Italy\\
%paolo.pistone@uniroma3.it}
%}

\maketitle

\begin{abstract}
In this paper we study a quantitative semantics for the $\lambda$-calculus based on tropical mathematics. Notably, we show that the relational model with values over the tropical semiring 
provides a unified framework for both 
program metrics, as based on the analysis of program sensitivity via Lipschitz-conditions, and for resource analysis, as based on higher-order program differentiation.
We discuss applications of this tropical semantics for ``best case'' resource analysis as well as for other quantitative properties like log-probabilities and differential privacy.
Finally, we show that a general foundation for this semantic approach is provided by an abstract correspondence between Lawvere's theory  
of generalized metric spaces and tropical algebra.

\end{abstract}

\begin{IEEEkeywords}
Differential lambda calculus, Tropical semiring, Lawvere quantale, Program metrics, Relational semantics.
\end{IEEEkeywords}

\section{Introduction}
\input{section1}

\section{A Bridge Between Two Quantitative Approaches to the $\lambda$-calculus}\label{section2}
\input{section2}

%\section{Tropical (non-)Linear Algebra and computation}
\subsection{Tropical linear algebra and the relational model% and computation
}

\input{section3intro}

%\subsection{As a model of the \emph{linear} $\STLC$}
\section{Tropical Semantics of Several $\lam$-Calculi}\label{section3}
\subsection{Prohibited duplication/erasure: the \emph{linear} $\STLC$}
\input{section3A}

%\subsection{As a model of the $\STLC$}
\subsection{Allowed duplication/erasure: the $\STLC$}
\input{section3B}

%\subsection{As a model of the $\STDLC$}
\subsection{Controlled duplication/erasure: the $\STDLC$}
\input{section3C}

\section{Tropical Laurent Series}\label{section4}
\input{section4}

\section{Applications}\label{sec:app}\label{section5}
\input{section5}
\input{section5C}

\section{Generalized Metric Spaces and Modules over the Lawvere Quantale}\label{section6}
\input{section6}%

\section{Related Work}\label{section7}
\input{section7}

\section{Conclusion and Future Work}\label{section8}
\input{section8}

\bibliographystyle{plain}
\bibliography{tropical.bib}

\onecolumn
\appendix

\subsection{Proofs from Section III}
\input{appIII}
\newpage

\subsection{Proofs from Section IV}
\input{appIV}

\newpage
\subsection{Proofs from Section V}
\input{appV}

\newpage
\subsection{Proofs from Section VII}
\input{appVII}

\end{document}