### 2023-03-18 rebuttal.txt
Committ pre-cena

We thank very much all the reviewers for the careful comments.
We will of course take care in correcting all the typos and unclear sentences.
Below we anwer to the main points.

Reviewer 1:

> p5 "setting $t^!(x) := t o_! x$"...

We are tacitly using the standard indentification of Q^X with the morphisms in QRel_!(T,X)=QRel(!T,X) where T=\emptyset is the terminal object.

>  p7 what is $[[M]]^!$?

It is the function discussed in the previous answer (i.e. in equation (6) at page 5).

> p7 "which can be written as a parameterized power series of the form" what does this mean? $f$ is a single function, but $f_t$ is an infinite family of functions. does it mean $f$ can be written as $f_t$ for some fixed constant $t$?"

We mean that we consider f a power series of shape \sum_n a_n x^n, and for fixed t>0 we take c_n := log_t (a_n), so that f can be written in the shape given in the paper. We agree that the subsection is not clear, so we will rephrase it better.

> p9 Prop 17: why is the input of $f$ suddenly $[0, 1]^X$?

We are supposing to see a database as a function X --> [0,1], which is indeed often the case (for instance, for a relational database -- take X as the key/attributes and [0,1] as the datas). We will mention that.

> Theorem 24 to bottom of column: I found this discussion hard to follow.

One of the core aspects of Theorem 24 is that on any L-module one can define a canonical metric; the goal of this paragraph is thus to describe in an explicit way the metrics associated with the SMCC structure of LMOD (or, equivalently, the metrics defining the SMCC structure of LCCAT). Maybe we should use M,N rather than X,Y in order to avoid confusion.  
Concerning tensors, in the previous page (after Prop. 19) we introduced the tensor product in LMOD and recalled that an element of it is a join of basic tensors x(x)y.

Reviewer 2:

First of all, we thank you for your thorough reading. We will surely adjust our presentation following your useful remarks!

> Comment 1, comment 2, comment 7, comment 8.

We were not aware of Lemay's "Coderelictions for Free Exponential Modalities" and we found its results very interesting/. As a general remark, our results on LREL and LREL_! in Section III are indeed based on relatively unsurprising checks; but since this paper is also addressed to the program metrics community, which is less familiar with linear logic and weighted relational models, this section is mainly thought to provide a relatively concrete presentation of (the tropical variant of) this model, before studying it in more detail in the following sections.
Therefore, our plan is to adjust Section III by making clearer what results are general (by also replacing the use of "Fact" with "Proposition/Lemma" toghether with the appropriate reference), and making clearer that the point of that section is to show how they look like in a tropical setting.
We would also accordingly adjust the appendix in case it should appear in an eventual final version of the paper.

> "Theorem 81 is important and should have appeared in the main paper."

Theorem 81, from the appendix, appears in the main text as Theorem 29, except for the mention of differential linear categories. We chose not to mention the latter in the main text because our results concerning them are either folklore or based on well-known results/techniques.

> "In the introduction, you say "the somehow canonical "Taylor semantics for STdLC". Why is that? What is the justification for this statement?"

We agree that the sentence is not clear, so will modify it. We simply meant that since Taylor and relational semantics are tightly related, it is natural to look at what happens in that kind of models.

Reviewer 3:

> "It is good that the model can characterize
    differential privacy of programs, but its implication to the
    verification tasks is not discussed."

You are right, the applications in Section 5 are just sketched. However, one of the main goals of this paper is to demonstrate that tropical methods, which are widely applied in computer science (but not in the linear logic/lambda-calculus community) can make sense also for studying the quantitative semantics of the lambda-calculus. Concrete developments of such applications are thus left for future work. We will clarify that point.

> "Another limitation of the
    paper's model is that it can not handle probabilistic programs
    involving continuous distributions"

This is indeed a well-known limitation of the semantics of probabilistic programs based on weighed relational models (e.g. probabilistic coherent spaces [33]). Still, such semantics are well-studied and led to strong results (e.g. full abstraction theorems). Anyway, we agree that this limitation should be mentioned.

> p.4 where is this hat operation defined?

It is supposed to be defined by those line. But we agree that it is not clear, so we will rephrase it better.

> "In complex analysis, Laurent series include terms of negative
    degrees, while here I think we only consider terms with
    non-negative degrees."

This is true, yet often in the literature on tropical mathematics (to which we refer, e.g. [59]) these are defined only with positive coefficients. We will explicitly mention it.

> "Is Theorem 24, which is one of the main claims in the paper,
    original, or subsumed by [71]?"

We think that the main contribution in Section 6 is to show that an abstract correspondence (expressed indeed by Theorem 24) known in the categorical literature can be extended to account for the relevant structure for interpreting the differential lambda-calculus. Hence, we see Theorems 27 and 29 as the main claims of this section, rather than Theorem 24 (whose statement and proof can be deduced from the results of [71]). We will clarify this.