### 2023-05-08 Notes/Maslov.tex
Ho creato anche questo branch per le altre note

\documentclass[11pt]{article}
%\providecommand{\event}{TLLA 2022} % Name of the event you are submitting to
\usepackage{breakurl}             % Not needed if you use pdflatex only.
\usepackage{underscore}           % Only needed if you use pdflatex.

%myPackages
\usepackage[T1]{fontenc} 
\usepackage[utf8]{inputenc} 
\usepackage{lmodern} 
\usepackage[english]{babel} 
\usepackage{amssymb}  
\usepackage{stmaryrd} 
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{ebproof}
\usepackage{tkz-graph}
\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary {positioning}
\usepackage{cmll}
\usepackage{mathrsfs}
\usepackage{marvosym}
\usepackage{cancel}
\usepackage{pst-solides3d}
\usepackage{MnSymbol}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing,calc}
%\usepackage[top=1in, bottom=1.25in, left=1in, right=1in]{geometry}

\setlength{\parindent}{0.7em}
\setlength{\parskip}{1.3em}

%\bibliographystyle{alpha}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{Lemma}[theorem]{Lemma}
\newtheorem{Proposition}[theorem]{Proposition}
\newtheorem{Corollary}[theorem]{Corollary}
\newtheorem{Definition}[theorem]{Definition}
\newtheorem{Example}[theorem]{Example}
\newtheorem{Remark}[theorem]{Remark}
\newtheorem{Notation}[theorem]{Notation}
\newtheorem{Terminology}[theorem]{Terminology}

\newcommand\eg{\textit{e.g.\ }}
\newcommand\etc{\textit{etc}}

\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{claim}{Claim}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{notation}{Notation}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{assumption}{Assumption}

\newcommand{\iu}{û}
\newcommand{\e}{ê}
\newcommand{\A}{â}
\newcommand{\Xplus}{X^{+}}
\newcommand{\forces}{\ \digamma}
\newcommand{\cons}{\textbf{.}}
\newcommand{\brabra}[1]{\llbracket #1 \rrbracket}
\newcommand{\test}[1]{\Vert #1 \Vert}
\newcommand{\str}[1]{\vert #1 \vert}
\newcommand{\sub}[3]{#1\{#2/#3\}}
\newcommand{\oppose}[3]{#1 \not\varepsilon #2^{+}(#3)}
\newcommand{\prog}[1]{\texttt{\small{\textbf{#1}}}}
\newcommand{\callcc}{\prog{callcc}}
\newcommand{\parts}{\mathcal{P}}
\newcommand{\fparts}{\parts_{f}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\rbeta}{\rightarrow_{\beta}}
\newcommand{\infer}[2]{\dfrac{#2}{#1}}
\newcommand{\interpret}[1]{\llbracket #1\rrbracket}
\newcommand{\Env}[1]{\mathrm{Env}_{#1}}
\newcommand{\restr}[2]{#1_{\big|_{#2}}}
\newcommand{\linc}{c^\bullet}

\usepackage{setspace}

% General Symbols
\newcommand{\nat}{\mathbb{N}} % natural numbers
\newcommand\seq\vec % sequence
\newcommand{\st}{s.t.\ } % Such that
\newcommand{\set}[1]{\{#1\}}
\newcommand{\bnfeq}{::=}
\newcommand\pow[1]{\mathscr{P}(#1)}  % powerset
\newcommand{\void}{\varnothing} % empty set
\newcommand{\fmsets}[1]{\mathcal{M}_f(#1)} % set of finite multisets on #1
\newcommand{\emptymset}{1} % empty multyset
\newcommand\bpf{\begin{proof}}
\newcommand\epf{\end{proof}}
\newcommand{\cT}{\mathcal{T}}
\newcommand{\cH}{\mathcal{H}}
\newcommand\cX{\mathcal{X}}
\newcommand{\finsub}{\subseteq
\newcommand{\cL}{\mathcal{L}}_\mathrm{f}}
% Lambda calculus
\newcommand{\lam}{\ensuremath{\lambda}}
\newcommand{\lamu}{\lambda\mu}
\newcommand{\Lam}{\ensuremath{\Lambda}}
\newcommand{\Lamo}{\Lambda^o}
\newcommand{\lamb}{\ensuremath{\lambda\bot}}
\newcommand{\subst}[2]{\{#2/#1\}} % substitution
\newcommand{\cB}{\mathcal{B}} % B of Bohm tree theory
\newcommand{\msto}[1][]{\twoheadrightarrow_{#1}} % multistep reduction
\newcommand{\holes}[1]{\langle #1\rangle} % context holes
\newcommand{\LamContext}[1]{\Lam\holes{#1}} % Lam<\xi>
\newcommand{\FV}[1]{\mathrm{FV}(#1)} % free variables
\newcommand{\BT}[1]{\mathrm{BT}(#1)} % BT(M)
\newcommand{\App}[1]{\mathcal{A}(#1)} % Approximants(M)
\newcommand{\Appr}{\mathcal{A}pp} % Set of all the approximants
\newcommand{\BTle}{\le_\bot} % <=_\Bot
\newcommand{\BTge}{\ge_\bot}
\newcommand{\comb}[1]{\mathsf{#1}}
\newcommand{\bI}{\comb{I}}
\newcommand{\rel}[1]{\mathtt{#1}}
\newcommand{\Om}{\comb{\Omega}}
\newcommand{\blam}{\boldsymbol{\lambda}}
\newcommand{\sqle}{\sqsubseteq}
\newcommand{\TM}[2][]{\mathcal{M}_{#2}^{#1}}
\newcommand{\fresh}{\mathrel{\#}}
\newcommand{\namedapp}[3]{(#1)_#2 #3}
% Resource calculus
\newcommand{\bag}[1]{{[}#1{]}}
\newcommand{\sums}[1]{\Bool\langle #1 \rangle} % 2<Lam^r>
\newcommand{\Bool}{\mathsf{2}} % ring 0,1
\newcommand{\Lamr}{\Lambda^r} % set of resource terms
\newcommand{\Lamb}{\Lambda^b} % set of bags
\newcommand{\lsubst}[2]{\langle [#2]/#1\rangle} % linear substitution
\newcommand{\Sum}[1]{\mathbb{#1}} % Sum of terms
\newcommand{\Te}[1]{\mathcal{T}(#1)} % (support of) Taylor expansion
\newcommand{\NF}[2][]{\mathrm{NF}_{#1}(#2)} % Normal form
\newcommand{\ResContext}[1]{\Lamr\holes{#1}} % Lam^r<\xi>
\newcommand{\BagResContext}[1]{\Lamb\holes{#1}} % Lam^b<\xi>
\newcommand{\NFTeq}{=_\tau} % "=_\tau", meaning NF(T(M)) = NF(T(M))
\newcommand{\size}[1]{\mid#1\mid} % size of a resource term
\newcommand{\nf}[2][]{\mathrm{nf}_{#1}(#2)}
\newcommand{\NFT}[1]{\mathrm{NF}\mathcal{T}(#1)}
\newcommand{\lin}[1]{#1^\circ}
\newcommand{\nak}[1]{|#1|}
\newcommand{\tor}{\to_r}
\newcommand{\redto}[1][]{\to_{#1}}
\newcommand{\emptybag}{1}
\newcommand{\forg}[1]{\lfloor #1\rfloor}
\newcommand\hole[1]{\llparenthesis\hspace{1pt} #1\hspace{1pt} \rrparenthesis} %context hole
\newcommand\linsub[2]{\langle #2 / #1 \rangle} %t<b/x> (lin sub without writing "[]" fo the bags)
\newcommand\lnamedapp[3]{\langle #1 \rangle_{#2} #3}
\newcommand\bigstep{\rightrightarrows}

%MACROS GIULIO
%!TEX root = ../BarbaM.tex
%%%%%% Lambda
\newcommand{\bnf}{::=}
\newcommand{\Appof}[1]{\mathcal{A}(#1)}
\newcommand{\approf}[1]{P_{#1}}
\newcommand{\bool}{\mathsf{2}}
\newcommand\Mfin[1]{\mathcal{M}_\mathrm{f}(#1)}
\newcommand{\cJ}{\mathcal{J}}
\newcommand{\cI}{\mathcal{I}}
\newcommand{\PLL}{\textsf{PLL}}
\newcommand\dg[2]{\mathrm{deg}_{#1}(#2)}
\newcommand{\comp}[1]{#1\!\!\uparrow}

\newcommand{\Pfinstar}[1]{\mathscr{P}_\mathrm{fin}^*(#1)}
\newcommand{\web}[1]{\lvert #1\lvert}
\newcommand{\homology}[2]{\mathcal{H}_{#1}(#2)}
\newcommand{\cs}[1]{\mathrm{cs}(#1)}
\newcommand{\simplices}[1]{\mathcal{S}#1}
\newcommand{\orsimpl}[2]{\overrightarrow{\mathcal S_{#1}} #2}
\newcommand{\barIexp}[2]{\mathscr{I}^{#1}\!#2}
\newcommand{\barI}[1]{\mathscr{I}\!#1}
\newcommand{\barsubd}[1]{\mathbf{bs}(#1)}

\newcommand{\lolli}{\multimap}
\newcommand{\cutelim}{\rightsquigarrow}

\newcommand{\minimum}[2]{\emph{min}\{#1,#2\}}

\newcommand{\alarm}[1]{\color{red}#1\color{black}}

\newcommand{\qRel}[1]{\mathrm{Rel_{#1}}}

\newcommand{\eps}{\epsilon}

\newcommand{\norm}[1]{\lVert #1 \lVert}
\newcommand{\supnorm}[1]{\lVert #1 \lVert_\infty}
\newcommand{\absv}[1]{\mid #1\mid}

\title{Notes on the Maslov dequantisation}
\author{Davide Barbarossa\\
Dipartimento di Informatica: Scienza ed Ingegneria\\
Università di Bologna%\\ Bologna, Italia\thanks{A fine university.}
\\
Bologna, Italia
}
%\def\titlerunning{Maslov dequantisation}
%\def\authorrunning{Davide Barbarossa}
\begin{document}

\maketitle

%\begin{abstract}
%\end{abstract}

Fix $t>0$.
Define $\phi_t:[0,1]\to[0,+\infty]$, $\phi_t(0<x):=-t\ln x$, $\phi_t(0):=+\infty$.
It is bijective with inverse $\phi^{-1}_t:[0,+\infty]\to[0,1]$, $\phi^{-1}_t(0<\alpha<+\infty):=e^{-\alpha/t}$, $\phi^{-1}_t(0):=1$, $\phi^{-1}_t(+\infty):=0$.

Call $\widetilde{\mathbb L}:=([0,1],\tilde+,\cdot)$, where $x\tilde+y:=(x+y)/2$.
This is a commutative non-associative idempotent semiring\footnote{Here by \emph{semiring} I mean a set endowed with an associative multiplication, an associative and commutative addition, with distributivity of the multiplication over addition. By \emph{non-associative semiring} I mean as before but without requiring associative addition. By \emph{idempotent} I mean with $x+x=x$.} with unit $1$ and pseudo-zero\footnote{For a non-associative semiring $R$, a \emph{pseudo-zero} $0$ is an element s.t.\ $x+0=f(x)=0+x$ and $x\cdot0=0=0\cdot x$, where $f$ is a fixed function $R\to R$ not depending on $x$.} $0$.

Call $\mathbb L_t:=([0,+\infty],+^t,\cdot^t)$, where $\alpha+^t\beta:=\phi_t(\phi_t^{-1}(\alpha)\tilde+\phi_t^{-1}(\beta))=-t\ln (e^{-\alpha/t}+e^{-\beta/t})+t\ln 2$,
and $\alpha\cdot^t\beta:=\phi_t(\phi_t^{-1}(\alpha)\phi_t^{-1}(\beta))=\alpha+\beta$.
This is a commutative non-associative idempotent semiring with unit $0$ and pseudo-zero $+\infty$.
$\mathbb L_1$ is usually called the \emph{unit-log-semiring}.

Remark that the bijection $\phi_t:\widetilde{\mathbb L}\to\mathbb L_t$ is an isomorphism of semirings with unit and pseudo-zeros, as $\phi_t(x\tilde+y)=\phi_t(x)+^t\phi_t(y)$ and $\phi_t(x\cdot y)=\phi_t(x)\cdot^t\phi_t(y)$.
Therefore, all the $\mathbb L_t$, for $t>0$, are isomorphic.

Call $\mathbb L:=([0,+\infty],\inf,+)$ the \emph{Lawvere tropical semiring}.
This is a commutative idempotent semiring with unit $0$ and zero $+\infty$.

Since $\mathbb L$ is associative, $\mathbb L_t$ and $\mathbb L$ cannot be isomorphic, for no $t>0$ (this can also be proven, without considering the associativity, by showing that there cannot exist a bijective homomorphism of semiring with units and pseudo-zeros between them).

It can be proven that $\alpha+^t\beta \underset{t\to 0}{\longrightarrow} \min\set{\alpha,\beta}$ (and trivially $\alpha\cdot^t\beta=\alpha+\beta \underset{t\to 0}{\longrightarrow} \alpha+\beta$).
In this sense, one says that $\mathbb L_t \underset{t\to 0}{\longrightarrow} \mathbb L$.
Looking at $\mathbb L$ as this limit is usually referred to as seeing $\mathbb L$ as the \emph{Maslov dequantisation} of $\mathbb{\widetilde L}$ (i.e.\ of the $\mathbb L_t$).

Remark that this limit deforms a \emph{isomorphic family} of commutative \emph{non-associative} idempotent semirings with units and \emph{pseudo-zeros}, to a \emph{non-isomorphic} commutative idempotent semiring with unit and \emph{zero}.

\end{document}