### 2023-05-15 TLLA23/BarbaPisto.tex
Committ

\documentclass[submission,%copyright,creativecommons
]{eptcs}
\providecommand{\event}{TLLA 2023} % Name of the event you are submitting to
%\usepackage{breakurl}             % Not needed if you use pdflatex only.
\usepackage{underscore}           % Only needed if you use pdflatex.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{bussproofs}
\EnableBpAbbreviations
\usepackage{hyperref}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
\usepackage{tikz-cd}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usetikzlibrary{cd}
%\usepackage{mathabx}
\usepackage{cmll}
\usepackage{stmaryrd}
\usepackage{todonotes}

\input{macros}

\title{Tropical Mathematics and the Lambda-Calculus}
\author{Davide Barbarossa
\institute{DISI, Universit\`a di Bologna}
\email{davide.barbarossa@unibo.it}
\and
\qquad\qquad Paolo Pistone
\institute{\qquad\qquad\qquad DISI, Universit\`a di Bologna}
\email{\qquad\qquad\qquad paolo.pistone2@unibo.it}
}
\def\titlerunning{Trends in Linear Logic and Applications}
\def\authorrunning{D.\ Barbarossa and P.\ Pistone}

\newcommand\eg{\textit{e.g.\ }}
\newcommand\etc{\textit{etc}}

\begin{document}
\maketitle

%\begin{abstract}We propose to study the interpretation of the $\lambda$-calculus in the framework of tropical mathematics, as a unified framework for both program metrics -- based on the analysis of program sensitivity via Lipschitz-conditions -- and for resource analysis -- based on higher-order program differentiation.We sketch the relation of this semantics to quantitative properties like differential privacy, convergence logprobabilities and Probabilistic Coherent spaces.Finally, we study the abstract correspondence between this tropical semantics and Lawvere’s generalised metric spaces.\end{abstract}

%\section*{Motivation}

In recent years, more and more interest in the programming language community has been directed towards the study of quantitative properties of programs like computing the number of computation steps or convergence probabilities, as opposed to purely qualitative properties like termination or program equivalence. 
In particular, two different quantitative approaches have received considerable attention from the programming language community.
On the one hand, the approach of \emph{program metrics} \cite{Reed2010, Gaboardi2017, Gabo2019} and \emph{quantitative equational theories} \cite{Plotk} is based on the observation that probabilistic or numerical algorithms are not thought to compute a target function $f$ \emph{exactly}, but only in an approximate way.
%hence, one is usually interested to know whether the computed function is \emph{not too far} from $f$, rather than equal to $f$.
This led to study denotational frameworks in which types are endowed with metrics measuring similarities in program behavior \cite{Reed2010},\cite{Bonchi2018}, \cite{Geoffroy2020, PistoneLICS, PistoneFSCD2022}.
On the other hand, there is the approach based on \emph{differential} \cite{difflambda}, \cite{difflambda}, \cite{Manzo2013, Breuvart2018, PistoneLICS2022} or \emph{resource-aware} \cite{Boudol1993} extensions of the $\lambda$-calculus, which is well-connected to the \emph{relational semantics}\cite{Manzo2012, Manzo2013, dill} and \emph{non-idempotent} intersection types \cite{decarvalho2018, Mazza2016}.
This led to study syntactic or denotational frameworks in which one can define a \emph{Taylor expansion} of programs.

In both approaches a crucial role is played by the notion of \emph{linearity}, in the sense of linear logic, i.e.~of using inputs exactly once. In metric semantics, linear programs correspond to \emph{non-expansive} functions, i.e.\ maps that do not increase distances; moreover, the possibility of duplicating inputs leads to interpret programs with a fixed duplication bound as \emph{Lipschitz-continuous} maps \cite{Gaboardi2017}.
By contrast, in the standard semantics of the differential $\lambda$-calculus, linear programs correspond to linear maps, in the usual algebraic sense, while the possibility of duplicating inputs gives rise to \emph{power series}.

The starting observation of this work is that, at a first glance, there seems to be a  ``logarithmic'' gap between the two approaches:
in metric models $n$ times duplication results in a $n$-Lipschitz \emph{linear} function $n\cdot x$, while in differential models this results in a non-Lipschitz \emph{polynomial} function $x^{n}$.
 At the same time, 
this gap may be overcome once we interpret these functions in the framework of tropical mathematics where, for instance, $x^{n}$ precisely reads as $n\cdot x$.

Tropical mathematics \cite{Simon} is a well established algebraic and geometrical framework, with tight connections with optimisation theory \cite{Sturmfelds}, where the usual ring structure of numbers based on addition and multiplication is replaced by the semiring structure given, respectively, by ``$\min$'' and ``$+$''.
For instance, the polynomial $p(x,y)=x^{2}+xy^{2}+y^{3}$, when interpreted over the tropical semiring, translates as the piecewise linear function
$
\trop f(x,y)=\min\{2x, x+2y, 3y\}
$.

A tropical variant of relational semantics has already been considered \cite{Manzo2013}, and shown capable of capturing \emph{best-case} quantitative properties.
Connections between tropical linear algebra and metric spaces have also been observed \cite{Fuji} within the abstract setting of \emph{quantale-enriched} categories \cite{Hofmann2014, Stubbe2014}.
However, a thorough investigation of the full power of the interpretation of the $\lambda$-calculus within tropical mathematics has not yet been undertaken.
We sketch here some first steps, with the idea of bridging the two approaches mentioned above, and suggesting the application of tropical methods to the study of the $\lambda$-calculus and its quantitative extensions.
This also scales to a 
more abstract level, leading to introduce a differential operator for continuous functors between \emph{generalized} metric spaces (in the sense of \cite{Lawvere1973}).
%We will only sketch some of our main results.% of our our analysis.

\section{The Tropical Semantics of Linear Logic }

\paragraph{Tropical mathematics in a nutshell}

We let the \emph{tropical semiring}  $\Lawv$, the structure at the heart of tropical mathematics, be $[0,\infty]$ with addition $\min$ and multiplication $+$.
This coincides with the \emph{Lawvere quantale} $\Lawv$ \cite{Hofmann2014, Stubbe2014}, i.e.\ $[0,\infty]$ with order $\geq$ and usual $+$ as the monoid action. 
$\Lawv$ is at the heart of the categorical study of metric spaces initiated by Lawvere \cite{Lawvere1973}, a viewpoint we will take in the last section.
A \emph{tropical polynomial} is a piece-wise linear function $\varphi:\Lawv\to \Lawv$ of the form $\varphi(x)=\min_{j=1,\dots,k}\{i_{j}x+\matr \varphi_{i_{j}}\}$, with $i_{j}\in\N$, $\matr \varphi_{i_{j}}\in\Lawv$ and $k$ finite.
Those are always Lipschitz functions.
For example, $\varphi_{n}(x)=\min_{i\leq n}\{ix+2^{-i}\}$, plotted in Fig~\ref{fig:plot1}.
A \emph{tropical root} of $\varphi$ is a point $x\in \Lawv$ where the minimum defining $\varphi$ is attained at least twice. %, i.e.\ where $\varphi$ is not differentiable, i.e.~where the slope of $\varphi$ changes.
E.g., the tropical roots of $\varphi_{n+1}$ are of the form $2^{-(i+1)}$, $i \leq n$.
%With this definition, tropical roots mimic the usual factorization property of roots: if $x_{0}$ is a root of $f$, this factorizes as $f(x)=\min\{x,x_{0}\}+ g(x)$. Yet, unlike in standard algebra, tropical roots can be computed in linear time \cite{Noferini2015}.
A \emph{tropical Laurent series} (of one variable $x\in\Lawv$), shortly a \emph{tLs}, is a function $\varphi:\Lawv\to \Lawv$ of the form $\varphi(x)=\inf_{n\in\N}\{nx+\matr \varphi_{n}\}$, with $\matr \varphi_{n}\in\Lawv$.
That is, a tLs is a ``limit'' of tropical polynomials of higher and higher degree.
For example $\varphi(x):=\inf_{i\in\N}\set{ix+2^{-i}}$ %that we will take as our running example, 
is the ``limit'' of the $\varphi_{n}$, see Fig~\ref{fig:plot1}. %Since $\inf$s are not in general $\min$s, the behavior of tLS may be less predictable than that of tropical polynomials.
%For instance, tropical roots for tLs (see \cite{Porzio2021}) may also include limit points.
Finally, for a polynomial/power series $f(x)=\sum_n a_nx^n$, one defines its \emph{tropicalization} $\trop f(\alpha):=\inf_n\set{-\log a_n+n\alpha}$.
TLs are in general not Lipschitz, and their study is less developed than that of tropical polynomials.

\paragraph{Tropical weighted relational semantics in a nutshell}

The study of matrices with values over the tropical semiring is a special case of the
\emph{weighted relational semantics} \cite{Manzo2013}, a well-studied semantics of the $\lambda$-calculus and linear logic:
for a fixed \emph{continuous} semi-ring $Q$, take the category $\QREL$ whose objects are sets and $\HOM \QREL X Y=Q^{X\times Y}$ (set-indexed matrices with coefficients in $Q$). %(\cite{Manzo2013} would call it $Q^\Pi$).
As expected, $Q^X$ is a $Q$-module and we can identify $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$.
Taking $Q:=\Lawv$ we obtain the \emph{tropical weighted relational model} $\LREL$.
%, where $\Lawv$ is the already introduced Lawvere quantale, seen as the idempotent complete semiring $(\BB R_{\geq0}\cup\set{\infty},\inf,\infty,+,0)$.
%The category $\LREL$ is well-defined because $\Lawv$ is a continuous semiring (w.r.t.\ its quantale order $\preceq$.
%This amounts to check that $\min$ and $+$ commute with the $\inf$ (as operations on $\BB R_{\geq0}\cup\set{\infty}$, which is immediate), and that $(\Lawv,\preceq)$ is a cpo with $\infty$ as bottom element (which is immediate since in $\Lawv$ we have $\vee = \inf$) .
Remark that the composition in $\LREL$ %is the tropicalisation of the one defining it in $\QREL$, i.e.
reads as \ $(s\circ t)_{a,c}:=\inf_{b\in Y}\set{s_{b,c}+t_{a,b}}$;
similarly, linear maps $f:\Lawv^X\to \Lawv^Y$ are of shape $f(x)_b=\inf_{a\in X} \set{x_a+\matr f_{a,b}}$, for some matrix $\matr f\in Q^{X\times Y}$ and are precisely those induced by $\matr f\in\HOM \LREL X Y$. We call them \emph{tropical linear}.
%\end{remark}
By applying known results (taken from \cite{Manzo2013},  {\color{red}CITARE LEMAYYYY}), one obtains that $\LREL$ gives rise to denotational models of several variants of the $\lam$-calculus:
first, $\LREL$ is a SMCC, i.e.\ a model of the linear $\STLC$.
Also, the coKleisli $\LREL_!$ is CCC, i.e.\ a model of $\STLC$, where $!$ is the usual multiset comonad (so $!X$ is the set of finite multisets on $X$).
Here, the coKleisli composition of $s\in\Lawv^{!Y\times Z}$ and $t\in\Lawv^{!X\times Y}$ is the matrix $s\circ_! t\in\Lawv^{!X\times Z}$ given by
$(s\circ_! t)_{\mu,c}:=
 \inf\limits_{n\in\N, b_1\dots,b_n\in Y, \mu = \mu_1+\cdots +\mu_n}
 \left\{s_{[b_1,\dots,b_n],c} + \sum_{i=1}^n t_{\mu_i,b_i}\right\}$.
 As usual, a matrix $t\in\HOM{\LREL_!}{X}{Y}$ yields a \emph{linear} map $\Lawv^{!X}\to\Lawv^Y$. However, we can also ``express it in the base $X$'' and see it as a \emph{non-linear} map $t^!:\Lawv^X\to\Lawv^Y$, by setting 
 $t^!(x):=t\circ_! x$.
%  (we are identifying $\Lawv^X$ with the set $\HOM{\LREL_!}{\emptyset}{X}$ of the \emph{points} of $X$).
% This is the notion of \emph{non-linear} map generated by the CCC-structure of $\LREL_!$.
 Concretely, we have $t^!(x)_b=\inf_{\mu\in !X} \set{\mu x+ t_{\mu,b}}$
 where $\mu x:=\sum_{a\in X} \mu(a)x_a$.
 These functions correspond then to generalised tLs, i.e.\ with possibly infinitely many variables (as many as the elements of $X$) and for $X=Y=\set{*}$, we get usual tLs of one variable.
Instead, if the support $\set{\mu\in!X\mid\matr f_{\mu,b}\neq\infty}$ of $f$ is \emph{finite}, we get %maps of shape $f(x)_b:=\min_{\mu\in\C F} \set{\mu x+ t_{\mu,b}}$, i.e.\ 
tropical polynomials in possibly infinitely many variables.
Furthermore, $!$ can be decomposed into a family of \emph{graded} exponentials $(!_n)_{n\in\N}$ turning $(\LREL,(!_n)_{n\in\N})$ in a model for a $\mathit{BLL}$-style simply typed $\lambda$-calculus, that we call $\BSTLC$ (see e.g.~\cite{Brunel2014}).

Finally, $\LREL_!$ can be equipped with a \emph{differential operator} $D:\HOM{\LREL}{!X}{Y}\to \HOM{\LREL}{!(X+X)}{Y}$ defined by: $(Dt)_{\mu\oplus\rho,b}:=t_{\rho+\mu,b}$ if $\#\mu=1$ and $:=\infty$ otherwise. %(where a multiset $\nu \in !(X+X)$ is identified with a disjoint sum of $\mu,\rho\in !X$).
This turns $(\LREL_!,D)$ into a CC$\partial$C, i.e.\ a model of the differential $\lambda$-calculus.
Remember that the (quantitative) Taylor expansion of an ordinary $\lambda$-term $M$ is an inductively defined series $\Te{M}$ of differential $\lambda$-terms, the only non-trivial case being $\Te{PQ} :=  \sum_{k=0}^{\infty} (\Der^{k}[P] \cdot Q^{k})0$.
It can be seen that the morphisms of $\LREL_!$ can always be Taylor expanded, and the series interpreting in $\LREL_!$ the Taylor expansion of a $\STLC$-term $M$, converges to the interpretation of $M$.

%
%
%The Taylor formula decomposes an unbounded application as a limit of bounded ones.
%So one might well ask whether this formula interprets a $\lambda$-term 
%as a limit of Lipschitz maps and thus bridging, in a sense, the metric and differential approaches previously mentioned.  
%%Here, a natural direction to look for is the \emph{relational semantics}, i.e.~the somehow canonical ``Taylor'' semantics for $\STDLC$. 
%But terms with bounded applications correspond to \emph{polynomials}, usually non-Lipschitz functions. 
%Yet, if these polynomials were tropical, the Taylor expansion would really decompose $\lambda$-terms via limits (indeed, $\inf$s) of Lipschitz maps: just like in mathematical analysis, unbounded application could be seen 
%as a limit of \emph{more and more sensitive} operations.
%This is precisely the content of our Corollary~\ref{cor:main}(3).

\section{Tropical Laurent Series and Metric Semantics for $\STLC$}

The main goal of this section is to show that the interpretation of $\STLC$ in the tropical relational model yields a metric semantics, where the spaces $\Lawv^{X}$ are endowed with the $\|\cdot\|_{\infty}$-norm metric, and programs are interpreted as \emph{locally Lipschitz} maps:
\begin{theorem}\label{cor:main}
For any $\lambda$-term $M$:
\begin{varenumerate}
\item if $\Gamma \vdash_{\BSTLC} M:A$, then $\model M^!:\model\Gamma \to \model A$ is a \emph{Lipschitz} map.
\item if $\Gamma \vdash_{\STLC}M:A$, then $\model M^!: \model\Gamma \to \model A$ is a \emph{locally Lipschitz} map.

Moreover, the Taylor expansion $\Te M$ decomposes $\model M^!$ into an $\inf$ of \emph{Lipschitz}~maps.
\end{varenumerate}
\end{theorem}
Recall that the syntactic Taylor expansion decomposes an unbounded application as a limit of bounded ones;
the result above lifts this decomposition to a semantic level, presenting a higher-order program as limits of Lipschitz maps: it provides thus a bridge between the metric and the differential approaches.

The proof of the result above passes through the study of topological and metric properties of tLs. 
We first studied tLs with respect to the topology induced by the $\|\cdot\|_{\infty}$-norm.
%
%In this section we study the properties of generalised tLs, from the point of view of mathematical analysis.
%In the following we will also refer to them as tLs.

% 
% We will call them simply \emph{tropical Laurent series (tLs)}.
% %Since in the general case of $\QREL$, $t^!$ is a Laurent series with operations in $Q$, let us call \emph{tropical Laurent series} the functions of shape $t^!$ for some $t\in\HOM{\LREL_!}{X}{Y}$.
%\end{remark}
%
%We find the usual notion of tLs of one variable as follows:

%\begin{remark}

\begin{proposition}\label{prop:nondecr+conc}
 Any tLs $f:\Lawv^X\to\Lawv^Y$ is non-decreasing and concave, w.r.t.\ the pointwise order, and continuous w.r.t.~the norm $\|\cdot\|_{\infty}$.
\end{proposition}

%
%The norm $\norm{\cdot}_\infty$ on $\BB R$ naturally induces a metric $\norm{x-y}_{\infty}$ over the spaces $\Lawv^{X}$.
%We will show that tLs satisfy suitable Lipschitz properties  w.r.t.~these metrics.
%\emph{Wlog}, we will often take $Y=\set{*}$.

Let us now consider metric properties. First, it can be seen that all tropical \emph{linear} functions $f: \Lawv^{X}\to \Lawv^{Y}$ are non-expansive. 
%\begin{proof}[Proof sketch]
%Using the fact that $f(\B x)_{b}= \inf_{a\in X}\matr f_{a,b}+\B x_{a}$,
%the problem reduces to checking that $|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})| = |\B x_{a}-\B y_{a}|\leq \| \B x-\B y\|_{\infty}$.\end{proof}
This result shows that, in analogy with that happens in usual metric semantics, linear programs are interpreted by non-expansive functions. 
%\begin{proof}
%Using $f(\B x)_{b}= \inf_{a\in X}\matr f_{a,b}+\B x_{a}$,
%first observe that $|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})| = |\B x_{a}-\B y_{a}|\leq \| \B x-\B y\|_{\infty}$; we now have
%$|f(\B x)_{b}-f(\B y)_{b}| \leq |(\inf_{a\in X}\matr f_{a,b}-\B x_{a})-(\inf_{a\in X}\matr f_{a,b}-\B y_{a})| \leq
%\sup_{a\in X}|(\matr f_{a,b}-\B x_{a})- (\matr f_{a,b}-\B y_{a})|\leq 
% \| \B x-\B y\|_{\infty}$.
%\end{proof}
More generally, linear maps with bounded exponentials yield Lipschitz maps:
\begin{proposition}\label{prop:boundedlip}
If a tLs $f: \Lawv^{X}\to \Lawv^{Y}$ arises from a bounded matrix $\matr f:!_{n}X\times Y\to \Lawv$, then $f$ is $n$-Lipschitz-continuous.
\end{proposition}
%\begin{proof}[Proof sketch]
%It follows from the above lines about tropical linear functions, plus the remark that, for all $x\in \Lawv^{X}$, $\norm{ !_{n} x-!_{n} y}_{\infty}\leq n\cdot \norm{ x- y}_{\infty}$, where $!_{n} x$ is the restriction of $! x$ to finite multisets of degree $\leq n$.%
%%Using the fact that $f(\B x)_{b}=\inf_{\mu\in \C M_{\leq n}(X)}\{ \matr f_{\mu,b}+ \mu (!_{n}\B x) \}$, where $!_{n}\B x\in \Lawv^{\C M_{\leq n}(X)}$ is given by 
%%$(!_{n}\B x)_{[a_{1},\dots, a_{k}]}=\sum_{i=1}^{k}\B x_{a_{i}}$, 
%%it suffices to check that $\| (!_{n}\B x)-(!_{n}\B y)\|_{\infty}\leq n\cdot \| \B x-\B y\|_{\infty}$ and apply Proposition \ref{prop:troplinear}.
%\end{proof}
This result is perfectly analogous to what happens in the metric models recalled in the introduction.%, the bounded exponentials $!_{n}$ playing the role of the re-scaling trick.
It also entails that any tropical polynomial $\varphi:\Lawv^{X}\to\Lawv$ is $\mathrm{deg}(\varphi)$-Lipschitz continuous.

\begin{wrapfigure}{r}{0.3\textwidth}%
%\begin{figure}
\begin{tikzpicture}[scale=0.55]
\begin{axis}[samples=250]
\addplot[yellow,domain=0:0.8] {1+0.02};

\addplot[orange,domain=0:0.8] {min(x+1/2, 1)+0.01};

\addplot[red,domain=0:0.8] {min(2*x+1/4, x+1/2, 1};
\addplot[blue,domain=0:0.8] {min(3*x+1/8,2*x+1/4, x+1/2, 1)-0.01};
\addplot[orange,domain=0:0.8] {min(4*x+1/16,3*x+1/8,2*x+1/4, x+1/2, 1)-0.02};

\addplot[violet,domain=0:0.8] {min(
10*x+1/1424,
9*x+1/712,
8*x+1/356,
7*x+1/128,
6*x+1/64,
5*x+1/32,
4*x+1/16,3*x+1/8,2*x+1/4, x+1/2, 1)-.03};

\end{axis}

\end{tikzpicture}
\caption{\small Plot of the tropical polynomials $\varphi_1,\varphi_2,\varphi_3,\varphi_4$ (from top to bottom), and of their limit tLs $\varphi$ (in violet).}
\label{fig:plot1}
%\end{figure} 
\end{wrapfigure} %%

Let us now %look at what happens with tLs, i.e.~when considering the full exponential $!$.
%First, we
first consider the interesting case of tLs with \emph{finitely many variables}. 
Let us start by an example, shown in Fig~\ref{fig:plot1}: the 1-variable tLs $\varphi(x)=\inf_{i}\{x+2^{-i}\}$ behaves \emph{locally} like the polynomials $\varphi_{n}(x)=\min_{i\leq n}\{x+2^{-i}\}$. 
%In particular, for all $\epsilon >0$, $f$ coincides on $[\epsilon,\infty]$ with some polynomial $\varphi_{n}$. More, precisely, $\varphi$ coincides with $\varphi_{n}$ for $\epsilon \geq 2^{-(n+1)}$, the smallest tropical root of $\varphi_{n}$.
However, at
%
%$x\in [\epsilon,\infty]$, with $\epsilon>0$
%It can be proven by hand that $\varphi(x)$ is a $\min$ for all $x>0$.
 $x=0$ we have that $\varphi(0)=\inf_{i\in\N} 2^{-i}=0$, and this is the only point where the $\inf$ is \emph{not} a $\min$.
Also, while the derivative of $\varphi$ is bounded on all $\BB R_{>0}$, at $x=0$ it tends to $\infty$.
This phenomenon is reminiscent of \cite[Example 7]{Ehrhard2005}.
%Differentials and Distances in Probabilistic Coherence Spaces. FSCD 2019 which actually motivated our first investigations.
In fact, these properties are shared by all tLs with {finitely} many variables, as shown by the following result (we identify $!\set{1,\dots,k}$ with $\N^k$, so the matrix of a tLs $f$ with variables $x=x_1,\dots,x_k$ can be given as a $\matr f:\N^k\to\Lawv$, and $f$ has shape $f(x)=\inf_{n\in \N^k}\set{nx+\matr f(n)}$, where $nx$ is the scalar product).

\begin{theorem}\label{theorem:fepsilon}
 Let $k\in\N$ and $f:\Lawv^k\to\Lawv$ a tLs with matrix $\matr f:\N^k\to\Lawv$.
 For all $0<\epsilon<\infty$, there is a \emph{finite} $\C F_\epsilon \subseteq \N^k$ such that 
% 
% \begin{enumerate}
%  \item If $\C F_\epsilon=\emptyset$ then $f=\infty$ on all $\Lawv^k$;
%  \item If $f(x_0)=\infty$ for some $x_0\in[0,\infty)^k$, then $\C F_\epsilon=\emptyset$;
  %\item 
$f$ coincides on all $[\epsilon,\infty]^k$ with the tropical {polynomial} $P_\epsilon(x):=\min_{n\in \C F_\epsilon}\set{nx+\matr f(n)}$.
% \end{enumerate}
\end{theorem}
%\begin{proof}[Proof sketch]
%Take $\C F_\epsilon$ the set of $n\in\N^{k}$ such that 
%$\widehat f(n)<\infty$ and $\widehat f(m)> \widehat f(n)+\epsilon$ holds for all $m\prec n$, where $\preceq$ is the pointwise order on $\N^k$.
%\end{proof}
The result above suggests that the local behavior of tLs with finitely many variables can be studied with the tools of tropical geometry (e.g.~tropical roots, Newton polygones). %as suggested in the next section. 
Moreover, a consequence of Theorem~\ref{theorem:fepsilon} is that all tLs with {finitely many} variables are always \emph{locally} Lipschitz on $\BB R_{>0}$.
The following result extends this property to all tLs, also covering the case with infinitely many variables.

\begin{theorem}\label{thmTLSlocLip}
 All tLs $\Lawv^X\to\Lawv$ are locally Lipschitz on $\BB R_{>0}^X$.
\end{theorem}
The core of the proof is a convex analysis argument showing that an arbitrary function $f:\Lawv^X\to\Lawv$ which is non-decreasing, concave and continuous, must be locally Lipschitz.

Finally, let us look at the differential operator $D$ of $\LREL_{!}$.
It translates into a differential operator $D_{!}$ turing a tLs $f:\Lawv^{X}\to \Lawv^{Y}$ into a tLs $D_{!}f:\Lawv^{X}\times \Lawv^{X}\to \Lawv^{Y}$, linear in its first variable, and given by $D_{!}f(x,y)_{b}=\inf_{a\in X, \mu\in !X}\left\{\matr f_{\mu+a}+x_{a}+\mu y\right\}$.
One can check that, when $f$ is a tropical polynomial, $D_{!}f$ gives the standard tropical derivative (see e.g.~\cite{Grigoriev2017}).
Finally, the Taylor formula enjoyed by $\LREL_!$ morphisms, yields a ``tropical'' Taylor formula for tLs of the form $f(x)=\inf_{n}\left\{D_{!}^{(n)}(f)(!_{n}x,\infty)\right\}$.

\section{Tropical Semantics and Quantitative Properties: Likelihoods of Reduction Paths}

Since algebraic and geometric properties in tropical mathematics are usually more tractable from a computational point of view, in several well-known applications (e.g.~for optimization problems related to machine learning \cite{Pachter2004, Zhang2018, Maragos2021}) one starts from a given modelled phenomenon, typically expressed by some polynomial function $f$, and studies 
what of its properties can be deduced from the \emph{tropicalization} $\trop f$ of $f$.
This idea suggests several natural directions in which the tropical semantics of a higher-order program could be used to deduce properties which can be expressed as an optimization problem.%: for a $\lambda$-term $M$, which can be expressed in $\LREL_!$ as a tropical polynomial or power series $\model M$, and we investigate what quantitative properties of $M$ can be deduced from it. In particular, it is useful to think of $\model M = \trop f$ for some polynomial or power series $f$.
We sketch here only one example, regarding probabilities, but we are currently working on other directions including \emph{differential privacy} and \emph{best case analysis}.

As a toy example, consider a probabilistic extension $\STLC_\oplus$ of $\STLC$, with a new ground type $\bool$, terms $\true,\false$ of type $\bool$, terms of shape $M\oplus_p N$ and $pM$, for $p\in[0,1]$, typed via the usual rules.
%\[ \dfrac{\Gamma\vdash M:A \qquad \Gamma\vdash N:A}{\Gamma\vdash M\oplus_p N:A}  \qquad\qquad \dfrac{\Gamma\vdash M:A}{\Gamma\vdash pM:A}\]
We add reduction rules:
$M\oplus_p N \to pM$ and $M\oplus_p N \to (1-p)N$,
so that $M\oplus_p N$ plays the role of a probabilistic coin toss of bias $p$.
$\STLC_\oplus$ can be seen as a fragment of the PCF in \cite{Manzo2013}.

Consider now the following term
$
 M:=(\true \oplus_p\false)\oplus_p((\true\oplus_p\false)\oplus_p(\false\oplus_p\true))
$
of type $\bool$.
Let us give addresses $\omega\in\set{00,01,100,101,110,111}$ to the occurrences of $\true,\false$ in $M$, reading $0$ as ``left'' and $1$ as ``right'' in the tree structure of $M$.
%The same addresses also represent all the different possible reduction paths from $M$ to a normal form.
Calling $q:=1-p$, there are the following six normal terms reachable from $M$:
$P_{00}(p,q)\true$, 
$P_{100}(p,q)\true$, 
$P_{111}(p,q)\true$, 
$P_{01}(p,q)\false$, 
$P_{110}(p,q)\false$,
$P_{101}(p,q)\false$,
where the $P$'s are the following monomials in $p,q$:
$P_{00}(p,q):=p^2$,
$P_{100}(p,q):=qp^2$,
$P_{111}(p,q):=q^3$,
$P_{01}(p,q):=pq$,
$P_{110}(p,q)=P_{101}(p,q):=q^2p$.
They correspond to the respective reduction path from $M$ to the normal term of the given address.
$P_{\omega}$ is then the probability (as a function of $p,q$) of obtaining the respective occurrence $\true_{\omega}$ or $\false_\omega$.%after all the coin tossings in $M$ are performed.
Thinking of $p,q$ as parameters, $P_{\omega}$ is the \emph{likelihood function} of the event $\omega$.
The polynomial $Q_{1}(p,q):=P_{00}(p,q)+P_{100}(p,q)+P_{111}(p,q)$ gives instead the whole probability of obtaining $\true$ after all the tossings (similarly a $Q_{0}(p,q)$ for $\false$).
This way, the probabilistic evaluation of $M$ is presented as a \emph{hidden Markov model} \cite{Baum1966}, a fundamental statistical model, and notably one to which tropical methods are generally applied \cite{Pachter2004}.
A typical question in this case would be: \emph{knowing that $M$ produced $\true$, which is the choice of the parameters $p,q$ that maximizes the probability that %the produced occurrence of $\true$ was a fixed $\true_{\omega_0}$ (i.e. that 
among the paths leading to $\true$, the one taken was a fixed $\omega_0$?}
%These are questions much in the style of those asked in {\color{red}Ref?} in the framework of {\color{red}??}.
Answering it, amounts at solving an optimisation problem related to $P_{\omega}, Q_\omega$, which is more easily solved via the tropicalizations $\trop P_{\omega},\trop Q_{\omega}$. 
We are looking for $p,q\in[0,1]$ s.t.\ $q=1-p$ and
$\max_{\omega\in\set{00,100,111}} \, P_\omega(x,y) = P_{\omega_0}(x,y)$.
This last condition is equivalent to ask that
$\min\limits_{\omega\in\set{00,100,111}} \, -\log P_\omega(x,y) = -\log P_{\omega_0}(x,y)$,
i.e.\
$(\trop Q_1)( -\log x, -\log y) = (\trop P_{\omega_0})( -\log x, -\log y)$.
%Passing through $\trop $ %P_\omega, \trop Q_\omega$ makes the problem easier, as this amounts to study tropical polynomials (for instance computing tropical roots can be done in linear time \cite{Noferini2015}). %the tropicalisation operator $\trop{}$ as well as the \emph{negative $\log$-probabilities} appear.

\begin{remark}
 By adapting \cite[Section IV]{Manzo2013}, it can be seen that $\LREL_!$ is a model of $\STLC_\oplus$.
 In particular, if we set $\model \bool:=\set{0,1}$, our running example $M$ is interpreted as $\model M\in\Lawv^{\set{0,1}}$ giving the tropicalised probabilities: $\model M_0=(\trop Q_0)(p,1-p)$, $\model M_1=(\trop Q_1)(p,1-p)$.
\end{remark}

%\begin{example}\label{ex:study}
For our $M$, we have $\trop Q_1(x,y)=\min\set{2x,y+2x,3y}$ and $\trop Q_0(x,y)=\min\set{x+y,2y+x}$.
Studying $\trop Q_1$, we see that $\trop Q_1(x,y)=3y$ iff $y\leq \frac{2}{3}x$, and it coincides with $2x$ otherwise.
Remembering that $3y=P_{111}(x,y)$, we can now solve our optimisation problem above for $\omega_0=111$:
via the substitution $x:=-\log p$, $y:=-\log (1-p)$, it is equivalent to $-\log (1-p)\leq -\frac{2}{3}\log p$, i.e.\ $1-p\geq p^{\frac{2}{3}}$.
This means that, for $p\in[0,1]$ s.t.\ $1-p\geq p^{\frac{2}{3}}$ (for example, $p=\frac{1}{4}$), the most likely occurrence of $\true$ to obtain, knowing that $M$ sampled $\true$ in its normal form, is $\true_{111}$.
Remembering that $2x=P_{00}(x,y)$, for the other values of $p$ (for example, $p=\frac{1}{2}$), the most likely $\true$ to be sampled is the occurrence $\true_{00}$.
We have thus answered our question.
%\end{example}
Also, the $p\in[0,1]$ s.t.\ $(p,1-p)$ is a root of $\trop Q_1$ or of $\trop Q_0$ provide the values of the bias of $\oplus_p$ for which there are at least two different \emph{equiprobable} paths from $M$ to its normal form.
Yet, we do not have a full understanding of the role of roots in this setting.
% Moreover, looking at Equation~\ref{eq:max}, we see that the $p\in[0,1]$ s.t.\ $(-\log p,-\log(1-p))$ is a tropical root of, say, $\trop Q_1$ are the values of the bias of $\oplus_p$ for which there are at least two different equiprobable occurrences of $\true$ that are smapled by $M$ during its tossings to normal form, knowing that some $\true$ was sampled.

\section{Getting rid of bases: $\Lawv$-modules as generalized metric spaces}

As we have seen, tropical semantics provides a viewpoint in which metric and differential properties are somehow unified.
This approach can be made more abstract, thanks to a fundamental categorical correspondence between tropical linear algebra (i.e.~the study of quantale modules over $\Lawv$) and the theory of Lawvere's \emph{generalized metric spaces} (see \cite{Fuji, Stubbe2006}).

A quantale module over $\Lawv$, shortly a \emph{$\Lawv$-module}%(not to be confused with a module on $\Lawv$ \emph{as a semiring})
, is a triple  $(M,\preceq, \star)$ where $(M, \preceq)$ is a sup-lattice, and $\star: \Lawv \times M \to M$ is a continuous (left-)action of $\Lawv$ on it, where continuous means that $\star$ commutes with both joins in $\Lawv$ and in $M$. 
%Notice that this defines the usual notion of module over the Lawvere quantale $\Lawv$, and this does \emph{not} coincide with the modules on $\Lawv$ \emph{as a semiring} (which are indeed well-studied and well-known to provide a $CC\partial C$ {\color{red}LEMAYYYY}). 
The most basic example of $\Lawv$-modules are given by the spaces $\Lawv^{X}$, with order and action defined pointwise.
We see them as modules given together with a fixed base, $X$.
Our Theorem~\ref{thm:no-base} says that one can actually give a semantic of high order programs without need to fix a base.

We already remarked that the tropical semiring coincides with the \emph{Lawvere quantale} $\Lawv$.
In particular, Lawvere was the first to observe that a (possibly $\infty$) metric on a set $X$ is nothing but a ``$\Lawv$-valued square matrix'' $d:X\times X\to \Lawv$ satisfying axioms like e.g.~the triangular law.
Indeed, such distance matrices correspond to $\Lawv$-\emph{enriched categories} (in short, a $\Lawv$-category) \cite{Lawvere1973, Hofmann2014, Stubbe2014}.
Many topological properties of metric spaces translate in this way into purely categorical ones. In particular, $\Lawv$-enriched categories which are \emph{cocomplete} in the (enriched) categorical sense (i.e.~all weighted colimits exist) satisfy usual metric completeness properties (e.g.~Cauchy-completeness or Isbell-completeness).

Now, any $\Lawv$-module $M$ can be endowed with a metric 
$M(x,y)=\inf\{\epsilon \mid \epsilon \star x \geq y\}$, yielding a cocomplete $\Lawv$-enriched category and, conversely, any cocomplete $\Lawv$-enriched category $X$ has a $\Lawv$-module structure with order $x\leq y$ when $X(y,x)=0$, and monoidal operation defined via a suitable weighted colimit.
This induces an isomorphism between the SMCC $\Mod$ of $\Lawv$-modules and their homomorphisms and the SMCC $\GMet$ of cocomplete $\Lawv$-enriched categories and cocontinuous functors (notice that functoriality in $\GMet$ is just non-expansiveness).

%
%A $\Lawv$-module homomorphism is a map $f:M\to N$ commuting with both joins and the $\Lawv$-action. We let $\Mod$ indicate the category of $\Lawv$-modules and their homomorphisms. 
%
%Observe that the category $\LREL$ is (equivalent to) a subcategory of $\Lawv\SF{Mod}$.
%While $\LREL$ corresponds to considering modules (the $\LREL^X$'s) whose vectors are given in coordinates w.r.t.\ a \emph{fixed base} (the set $X$), $\Lawv\SF{Mod}$ corresponds to considering modules in abstract, without fixing a base.

%Indeed, spelling out the definition, a $\Lawv$-enriched category is given by a set $X$ together with a ``hom-set'' $X(-,-):X\times X\to \Lawv$ satisfying 
%\begin{align}
%0  & \geq X(x,x) \tag{$\Lawv$-cat 1}\\
%X(y,z)+X(x,y)&\geq  X(x,z) \tag{$\Lawv$-cat 2}
%\end{align}
%This structure is often referred to as a \emph{generalized metric space} \cite{Lawvere1973, Hofmann2014, Stubbe2014}. %, or as a \emph{quasi-metric space}. 
%A $\Lawv$-category is \emph{skeletal} \cite{Stubbe2014} when $X(x,y)=0$ implies $x=y$, and \emph{symmetric} when it coincides with its opposite category $X^{\op}(x,y):=X(y,x)$, i.e.~when $X(x,y)=X(y,x)$. 
%A metric space, in the usual sense, is just a particular case of $\Lawv$-category.
%%Notice that any $\Lawv$-category $X$ induces a skeletal category $X^{\sk}$ by quotienting points under $X(x,y)=0$, and a symmetric one by letting $X^{\sym}(x,y)=\max\{X(x,y),X^{\op}(x,y)\}$.
%Remark that $\Lawv$ has a canonical $\Lawv$-enriched structure given by $\Lawv(r,s)=s \overset{\cdot}{-} r$ (where ``$ \overset{\cdot}{-}$'' indicates truncated subtraction), 
%and the Euclidean distance coincides with its symmetrization $\Lawv^{\sym}(x,y):=\max\{\Lawv(x,y),\Lawv^{\op}(x,y)\}$.

%A $\Lawv$-enriched \emph{functor} between $\Lawv$-categories $X,Y$ is just a non-expansive map $f:X\to Y$, as functoriality reads as $Y(f(x),f(y))\leq X(x,y)$.
%Now, by considering distributors, one defines the notion of \emph{continuous} functor $f:X\to Y$ as one that commutes with all existing weighted colimits in $X$.
%A $\Lawv$-enriched category $X$ is said \emph{categorically complete} (or just \emph{complete}) if all weighted colimits over $X$ exist. 
%We let $\GMet$ indicate the category of complete and skeletal $\Lawv$-categories and continuous functors.
%We mention this notions because it can be shown (~\cite{Stubbe2006}) that, actually, $\Mod$ and $\GMet$ are isomorphic categories.

Our contribution was, first, to observe that the SMCC structure of $\LREL$ lifts to that of the more general category $\Mod\equiv\GMet$; secondly, to define an exponential $!$ in $\Mod\equiv\GMet$, using a well-known recipe from \cite{Mellies2018, Manzo2013} based on the construction of symmetric algebras, and show that in this way also the exponential and differential structure of $\LREL_{!}$ lift to this more general setting:
%
%This is interesting for us because it suggests to transport the metric point of view of $\GMet$ onto our modules in $\Mod$.
%Actually, we can even show that:
\begin{theorem}\label{thm:no-base}
$\Mod_{!}$ or, equivalently, $\GMet_{!}$, can be equipped with a differential $D$ making it a $CC\partial C$.
\end{theorem}
%In order to ,, consisting in defining a symmetric algebra %$\Sym_{n}(M)$
%as the equalizer of all permutative actions on $n$-tensors $M\otimes \dots \otimes M$, and then define $!$ as the product of the symmetric algebras.

%The optional arguments of {\tt $\backslash$documentclass$\{$eptcs$\}$} are
%\begin{itemize}
%\item at most one of
%{\tt adraft},
%{\tt submission} or
%{\tt preliminary},
%\item at most one of {\tt publicdomain} or {\tt copyright},
%\item and optionally {\tt creativecommons},
%  \begin{itemize}
%  \item possibly augmented with
%    \begin{itemize}
%    \item {\tt noderivs}
%    \item or {\tt sharealike},
%    \end{itemize}
%  \item and possibly augmented with {\tt noncommercial}.
% \end{itemize}
%\end{itemize}
%We use {\tt adraft} rather than {\tt draft} so as not to confuse hyperref.
%The style-file option {\tt submission} is for papers that are
%submitted to {\tt $\backslash$event}, where the value of the latter is
%to be filled in in line 2 of the tex-file. Use {\tt preliminary} only
%for papers that are accepted but not yet published. The final version
%of your paper that is to be uploaded at the EPTCS website should have
%none of these style-file options.

%By means of the style-file option
%\href{http://creativecommons.org/licenses/}{creativecommons}
%authors equip their paper with a Creative Commons license that allows
%everyone to copy, distribute, display, and perform their copyrighted
%work and derivative works based upon it, but only if they give credit
%the way you request. By invoking the additional style-file option {\tt
%noderivs} you let others copy, distribute, display, and perform only
%verbatim copies of your work, but not derivative works based upon
%it. Alternatively, the {\tt sharealike} option allows others to
%distribute derivative works only under a license identical to the
%license that governs your work. Finally, you can invoke the option
%{\tt noncommercial} that let others copy, distribute, display, and
%perform your work and derivative works based upon it for
%noncommercial purposes only.

%Authors' (multiple) affiliations and emails use the commands
%{\tt $\backslash$institute} and {\tt $\backslash$email}.
%Both are optional.
%Authors should moreover supply
%{\tt $\backslash$titlerunning} and {\tt $\backslash$authorrunning},
%and in case the copyrightholders are not the authors also
%{\tt $\backslash$copyrightholders}.
%As illustrated above, heuristic solutions may be called for to share
%affiliations. Authors may apply their own creativity here \cite{multipleauthors}.

%\section*{Bibliography}

%We request that you use
%\href{http://eptcs.web.cse.unsw.edu.au/eptcs.bst}
%{\tt $\backslash$bibliographystyle$\{$eptcs$\}$}
%\cite{bibliographystylewebpage}, or one of its variants
%\href{http://eptcs.web.cse.unsw.edu.au/eptcsalpha.bst}{eptcsalpha},
%\href{http://eptcs.web.cse.unsw.edu.au/eptcsini.bst}{eptcsini} or
%\href{http://eptcs.web.cse.unsw.edu.au/eptcsalphaini.bst}{eptcsalphaini}
%\cite{bibliographystylewebpage}. Compared to the original {\LaTeX}
%{\tt $\backslash$biblio\-graphystyle$\{$plain$\}$},
%it ignores the field {\tt month}, and uses the extra bibtex fields {\tt eid}, {\tt doi}, {\tt ee} and {\tt url}. The first is for electronic identifiers (typically the number $n$ indicating the $n^{\rm th}$ paper in an issue) of papers in electronic journals that do not use page numbers. The other three are to refer, with life links, to electronic incarnations of the paper.

%\nocite{*}
%\bibliographystyle{eptcs}
%\bibliography{generic}

\bibliographystyle{plain}
\bibliography{tropical.bib}

\end{document}