### 2023-07-27 CSL24/section5.tex

%%In this section we illustrate a few directions in which the tropical semantics just introduced could be used to analyze quantitative properties of higher-order programs.

%Since algebraic and geometric properties in tropical mathematics are usually more tractable from a computational point of view, in several well-known applications (e.g.~for optimization problems related to machine learning \cite{Pachter2004, Zhang2018, Maragos2021}) one starts from a given model, typically expressed by some polynomial function $f$, and studies  what properties of the model can be deduced from the \emph{tropicalization} of $f$, noted $\trop f$, i.e.~the transformation of $f$ into a tropical polynomial. Here we follow a similar pattern: we consider a program $M$, which can be expressed in the form of a polynomial or a power series $f$, and we  investigate what quantitative properties of $M$ can be deduced from the properties of $\trop f$, that will indeed coincide with the interpretation of $M$ in $\LREL_{!}$.

%
%
%%several well-known applications of tropical mathematics is to study how much can be deduced of some function starting from the properties of its tropicalization.
%%In Section \ref{section5} we will follow a similar direction, investigating what quantitative properties of a higher-order programs are revealed by the study of its tropical interpretation.
%

%\input{tropicalization}

\begin{comment}

\subsection{Best case analysis and metric reasoning}

The possibility of using the relational model over the tropical semiring for ``best case'' resource analysis has already been explored in \cite{Manzo2013}. Notably, they considered an interpretation of a language for $\B{PCF}$ with non-deterministic choice in which each $\lambda$-abstraction and each occurrence of the fixpoint operator $Y$ is assigned a ``weight'' 1, and showed that for any program $M$ of type $\B{nat}$, 
the value of the interpretation $\model{M}\in \Lawv^{\BB N}$ on a particular natural number $k$, i.e.~$\model{M}(k)\in \Lawv$, corresponds to the \emph{minimum} number of $\beta$- or $\TT{fix}$-redexes reduced in a reductions sequence from $M$ to $\underline n$. 
In the next paragraph we will illustrate an analogous ``best case'' analysis for probabilistic programs.

What does the metric analysis from the previous sections add to that? Firstly, the possibility of \emph{comparing} different programs with respect to their quantitative properties. For example, in the $\B{PCF}$ semantics recalled above, the distance between two programs $M$ and $N$ of type $\B{nat}$, provides a bound on the difference between the  ``best case'' computation time of $M$ and that of $N$. For instance, by taking, instead of the $\infty$-norm metric on $\Lawv^{\BB N}$,  
the \emph{non symmetric} distance (or quasi-metric, a viewpoint we explicitly take in Section \ref{section6}) $q(\B x, \B y)=\sup_{n}\{\B y_{n}\dotdiv \B x_{n}\}$, a ``distance'' $q(\model{M},\model{N})\leq \epsilon$ would indicate that $\model{M}$ \emph{improves} on $\model{N}$ of at most $\epsilon$ steps at each computation.

Secondly, the Lipschitz conditions from Section \ref{section4} allow us to reason on program distances in a \emph{compositional} way: suppose, as before, that $M,N:A$ are two programs such that $M$ improves on $N$ by $\epsilon$, and let $\TT C[-]:A \to \B{nat}$ indicate a context; knowing that the interpretation of $\TT C$ is $k$-Lipschitz-continuous on some open set containing both $\model M$ and $\model N$, allows us to immediately deduce that $\TT C[M]$ improves on $\TT C[N]$ by $k \epsilon$. 
Observe that this will typically be the case when the Taylor expansions of $\TT C[M]$ and $\TT C[N]$ actually yields a \emph{finite} sum of at most $k$ terms, i.e.~when 
\begin{align}
\TT C[M] = \sum_{i=0}^{k} \TT D^{(k)}\Big[\lambda x.\TT C[x]\Big]\cdot M^{k}
\end{align}
and similarly for $\TT C[N]$. It is here worth recalling that, for $\STDLC$, a well-known result \cite{difflambda} is that the Taylor expansion of a closed application $MN$ is always \emph{finite}, although its non-zero coefficients may be arbitrarily high. 
Notably, these observations suggest to study tropical versions of \emph{finiteness spaces} \cite{Ehrhard2005}, 
a variant of the relational semantics modeling strongly normalizing programs via \emph{finite} power series.
%We mention this point in Section~\ref{section8}.

\end{comment}

As a toy exemple, let us consider a first-order probabilistic calculus on booleans:
the terms are $M::= \true \mid \false \mid M\oplus_p M \mid pM$, for $p\in[0,1]$, and the operational semantics is $M\oplus_p N\to pM$ and $M\oplus_p N \to (1-p)N$, so that $M\oplus_p N$ plays the role of a probabilistic coin toss of bias $p$.

Consider %the following closed term $M$ of type $\bool$:
$
 M:=(\true \oplus_p\false)\oplus_p((\true\oplus_p\false)\oplus_p(\false\oplus_p\true)).
 $
Let us give addresses $\omega\in\set{ll,lr,rll,rlr,rrl,rrr}$ to the occurrences of $\true,\false$ in $M$, by following the tree structure of $M$, ($l$ is ``left'' and $r$ is ``right'').
The same addresses also represent all the different possible reduction paths from $M$ to a normal form.
%For instance, $rll$ represents the reduction which keeps the right part of the outermost $\oplus_p$ and erases the left part, then continues by choosing the left part twice, reaching at the end the occurrence $\true_{rll}$ in $M$, i.e.\ the second occurrence of $\true$ in $M$ starting from the left.
Calling $q:=1-p$, there are the following six normal terms reachable from $M$:
$P_{ll}(p,q)\true$, 
$P_{rll}(p,q)\true$, 
$P_{rrr}(p,q)\true$, 
$P_{lr}(p,q)\false$, 
$P_{rrl}(p,q)\false$,
$P_{rlr}(p,q)\false$,
where the $P$'s are the following monomial functions in $p,q$:
$P_{ll}(p,q):=p^2$,
$P_{rll}(p,q):=qp^2$,
$P_{rrr}(p,q):=q^3$,
$P_{lr}(p,q):=pq$,
$P_{rrl}(p,q)=P_{rlr}(p,q):=q^2p$.
%They correspond to the respective reduction path from $M$ to the normal term of the same address.
$P_{\omega}(p,1-p)$ is then the probability of the event ``$M\twoheadrightarrow \true_\omega/\false_\omega$'' (depending on $\omega$).
Thinking of $p,q$ as parameters, $P_{\omega}(p,q)$ can thus be read as the \emph{likelihood function} of the event ``$M\twoheadrightarrow \true_\omega/\false_\omega$''.
The polynomial function $Q_{\true}(p,q):=P_{ll}(p,q)+P_{rll}(p,q)+P_{rrr}(p,q)=p^2+p^2q+q^3$ gives instead the probability of the event ``$M\twoheadrightarrow \true$'', and analogously for $Q_{\false}(p,q):=P_{lr}(p,q)+P_{rrl}(p,q)+P_{rlr}(p,q)=pq+2pq^2$.
%Let us consider in this subsection a probabilistic extension of $\lam$-calculus, call it $\STLC_\oplus$, adding as usual terms of shape $pM+qN$ and $M\oplus_p N$, for $p,q\in[l,r]$.These terms are typed via the rule:
%\[
%\dfrac{\Gamma\vdash M:A \qquad \Gamma\vdash N:A}{\Gamma\vdash M\oplus_p N:A}
%\]
%and similar for $\Gamma\vdash pM+qN:A$.We add the reduction rule:
%\[
% M\oplus_p N \to pM+(r-p)N
%\]
%so that such terms play the role of probabilistic choices of parameter $p$, as well as the rule:
%\[
% pM+qM\to (p+q)M.
%\]
%Let us consider $M:=(I\oplus_p\Omega)\oplus_p((I\oplus_p\Omega)\oplus_p(I\oplus_p\Omega))$. Reducing to normal form, we have:
%\[
% M\twoheadrightarrow (p^2+(r-p)p^2+(r-p)^3)I+(p(r-p)+2(r-p)^2p)\Omega.
%\]
%The index $\omega\in\set{ll,rll,rrr,lr,rrl,rlr}$ of each $P_\omega$ indicates the path of the reduction that led from $M$ to the respective occurrence $I_\omega$ of $I$ or $\Omega_\omega$ of $\Omega$ from $M$ to its normal form ($l$ means ``left'' and $r$ means ``right'').For instance, in order to reach $I_{rll}$, i.e.\ the second occurrence of $I$ from the left in $M$, we have to take the right path in the outer $\oplus_p$ of $M$, then two times the left path in the new outer $\oplus_p$'s that we encounter during reduction.$P_{\omega}(p,q)$ gives then the probability (as a function of $p,q$) of obtaining the respective occurrence $I_{\omega}$ or $\Omega_\omega$ in the normal form, if we were to sample at each time we reduce a $\oplus_p$.It can thus also be read as the likelihood function of such event.The polynomials $Q_{r,2}(p,q)$ give instead the whole probability of obtaining respectively $I$ or $\Omega$, in the normal form after such samplings.
This way, the probabilistic evaluation of $M$ is presented as a \emph{hidden Markov model} \cite{Baumr966}, a fundamental statistical model, and notably one to which tropical methods are generally applied \cite{Pachter2ll4}.

Typical questions in this case would be, for a fixed $\omega_0$:
%
%The tropical point of view allows now to express two natural questions about this situation:
\begin{enumerate}
 \item Which is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow \false_{\omega_0}$''?
 I.e., which is the choice of $p,q$ that maximizes the probability $P_{\omega_0}$ of the event ``$M\twoheadrightarrow \false_{\omega_0}$''  ?
 \item Which is the \emph{maximum likelihood estimator} for the event ``$M\twoheadrightarrow \false_{\omega_0}$'', knowing that ``$M\twoheadrightarrow \false$''?
I.e., which is the choice of $p,q$ that makes $\omega_0$ the most likely path among those leading to $\false$ (i.e.\ that maximizes the conditional probability $\BB P(``M\twoheadrightarrow \false_{\omega_0}'' \mid ``M\twoheadrightarrow \false'')$)?
\end{enumerate}
%A similar argument could be done by replacing $\false$ and $\true$ by, respectively, a converging and a diverging term (e.g.~in a $\B{PCF}$-style language), so r) would be about finding maximum likelihood estimators for the event ``$M$ converges''.

Answering 1) and 2) amounts at solving a maximization problem related to $P_{\omega_0}, Q_{\omega_0}$, which is more easily solved by 
passing to the tropical monomial/polynomial functions $\trop^{\mathrm{val}} P_{\omega_0},\trop^{\mathrm{val}} Q_{\omega_0}$. 
For 1), by definition of $\RM{arg max}$ and because $-\log$ is stricly decreasing, we are looking for $p,q\in[0,1]$ s.t.\ $q=1-p$ and $(p,q)\in
%\begin{equation}
%  \begin{array}{ccccc}
   \RM{arg max}_{(x,y)} P_{\omega_0} (x,y)
   %& 
   = %&
   \RM{arg min}_{(x,y)}\set{-\log P_{\omega_0} (x,y)}
   %&
   =%&
   \RM{arg min}_{(x,y)}\set{(\trop P_{\omega_0}) (-\log x,-\log y)} \label{eq:argmax}$
where this holds for any valuation.
Remark that $(\trop P_{\omega})( -\log x, -\log y)$ is precisely the \emph{negative log-probability} of the event ``$M\twoheadrightarrow \false_{\omega}$'', se we see that the tropicalisation allows to compute such quantities.
%  \end{array}
%\end{equation}
For 2), %the $\omega_l$ is s.t.\ $\max\limits_{\omega\in\set{ll,rll,rrr}} \, P_\omega(x,y) = P_{\omega_l}(x,y)$. So
we are looking for $p,q\in[0,1]$ s.t.\ $q=1-p$ and
$\max_{\omega\in\set{lr,rrl,rlr}} \, P_\omega(p,q) = P_{\omega_0}(p,q)$, i.e.\ $\min_{\omega\in\set{lr,rrl,rlr}} \, -\log P_\omega(p,q) = -\log P_{\omega_0}(p,q)$.
Remembering that $\trop^{\mathrm{val}} Q_{\false}(p,q)=\min\set{p+q, \mathrm{val}(2)+p+2q}$, we see that our minimization problem is equivalent to the equality $(\trop^0 Q_{\false})( -\log p, -\log q) = (\trop^0 P_{\omega_0})( -\log x, -\log y)\label{eq:max}$
and this holds only for the null valuation.
Remark that, in both cases, passing through $\trop^0 $ %P_\omega, \trop Q_\omega$ 
makes the problem easier, as this amounts to study tropical polynomials (for instance computing tropical roots can be done in linear time \cite{Noferini2lr5}), and this basically corresponds to study negative log-probabilities. %the tropicalisation operator $\trop{}$ as well as the \emph{negative $\log$-probabilities} appear.

%For our running example $M$, we have $\trop Q_{\true}(x,y)=\min\set{2x,y+2x,3y}$ and $\trop Q_{\false}(x,y)=\min\set{x+y,2y+x}$. Studying $\trop Q_{\true}$ %, whose plot is in Fig.~\ref{fig:plot2}, we see that $\trop Q_{\true}(x,y)=3y$ iff $y\leq \frac{2}{3}x$, and it coincides with $2x$ otherwise. Remembering that $3y=P_{rrr}(x,y)$, we can now solve the optimisation problem~\ref{eq:max} for $\omega_l=rrr$: via the substitution $x:=-\log p$, $y:=-\log (r-p)$, Equation~\ref{eq:max} is equivalent to $-\log (r-p)\leq -\frac{2}{3}\log_c p$, i.e.\ $r-p\geq p^{\frac{2}{3}}$. This means that, for $p\in[0,1]$ s.t.\ $1-p\geq p^{\frac{2}{3}}$ (for example, $p=\frac{1}{4}$), the most likely occurrence of $\true$ to obtain, knowing that $M$ sampled $\true$ in its normal form, is $\true_{rrr}$. Remembering that $2x=P_{ll}(x,y)$, for the other values of $p$ (for example, $p=\frac{1}{2}$), the most likely $\true$ to be sampled is the occurrence $\true_{ll}$. We have thus answered question 2) above for $\true$.

From this situation we notice the following important:

\begin{remark}\label{rmk:tropof01Rel}
This toy first-order language can be interpreted in the already mentioned $\overline{R_{\geq 0}}\mathrm{Rel}$ and in $\LREL$.
We do not give details now since, from the following section, we will introduce the interpretation of interesting high-order calculi, including a probabilistic one containing the one of this section.
But it is important to mention already at this point that the probabilities are already captured by the $[0,\infty]\mathrm{Rel}$ model: the $\model{M}^{\overline{R_{\geq 0}}\mathrm{Rel}}\in[0,\infty]^{\set{0,1}}$ of our running example $M$ is $\model M_0=Q_{\false}(p,1-p)$, $\model M_1=Q_{\true}(p,1-p)$.
Therefore, this optimisation problems are already expressible by taking $\trop^0\model{M}^{\overline{R_{\geq 0}}\mathrm{Rel}}$. 
Now, the model $\LREL$ is precisely null-valuation tropicalisation of $\overline{R_{\geq 0}}\mathrm{Rel}$, i.e.\ $\model{M}^{\LREL}=\trop^0\model{M}^{\overline{R_{\geq 0}}\mathrm{Rel}}$.
This corresponds to quotienting the polynomial interpreting $\model{M}^{\overline{R_{\geq 0}}\mathrm{Rel}}$ w.r.t.\ idempotent sum, as the null valuation eliminates all the coefficients (different from $1$).
A precise study of the ``tropicalisation of $\overline{R_{\geq 0}}\mathrm{Rel}$'' is left for future investigations, as it is related with power series arising from more sophisticated calculi with paramethers (in the style of \cite{} {\color{red}Ehrhard!}).
We therefore have that $\model{M}^{\LREL}_0(-\log p,-\log (1-p)$ is the negative log-probability of \emph{any} of the (equiprobable) most likely paths of reduction to normal form.
We take these observations as a motivation for considering such model, and the point of this work is to dig into such model, concentrating for this first paper on the relations between the metric and differential tools that we will consider.
% (in $p$ and $q$, not after the substitution $q:=r-p$) 
% are extracted from it, \ref{eq:argmax}, \ref{eq:max} of $\STLC_\oplus$-programs. 
\end{remark}

%Now, in $\LREL$, seen as a model of such probabilistic $\lam$-calculus, the interpretation of a term already computes the tropicalisation of the polynomials expressing the probabilities, because the underlying semiring of the model is tropical.
%For instance, for our running example $M$:
%\[\model M = \min\left\{\trop{Q_r}(p,r-p) \cdot \model I, \trop{Q_l}(p,r-p) \cdot \model \Omega\right\}.\]