### 2023-07-28 CSL24/main.tex
Altro commit pre-cena

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{bussproofs}
\EnableBpAbbreviations
\usepackage{hyperref}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
\usepackage{tikz-cd}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usetikzlibrary{cd}
\usepackage{mathabx}
\usepackage{cmll}
\usepackage{stmaryrd}
\usepackage{wrapfig}
\usepackage{relsize}

\input{prooftree}

\usepackage{proof}

\bibliographystyle{plainurl}% the mandatory bibstyle

\input{macros}

\title{Tropical mathematics for the Lambda-calculus 1}
\subtitle{The potential of tropical wheighted relational semantics%for probabilities and Taylor expansion
} %TODO Please add

\titlerunning{Tropical mathematics for the Lambda-calculus 1} %TODO optional, please use if title is longer than one line

\author{Davide {Barbarossa}}{Dipartimento di Informatica, Universit\`a di Bologna, %[optional: Address], 
Italy% \and My second affiliation, Country 
\and \url{https://lipn.univ-paris13.fr/\~barbarossa/index.html} }{davide.barbarossa@unibo.it}{https://orcid.org/0000-0003-4608-8282}{%(Optional)Funding each author
}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Paolo {Pistone}}{Dipartimento di Informatica, Universit\`a di Bologna, %[optional: Address], 
Italy% \and LIP, Universit\'e de Lyon, France 
\and \url{%WebPage
} }{paolo.pistone2@unibo.it}{%Orchid
}{%(Optional)Funding each author
}

\authorrunning{D.\,Barbarossa and P.\,Pistone} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Davide Barbarossa and Paolo Pistone} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[500]{Theory of computation~Lambda calculus}
\ccsdesc[300]{Theory of computation~Categorical semantics}
\ccsdesc[100]{Theory of computation~Linear logic}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Differential lambda-calculus, Tropical semiring, Relational semantics, Lawvere quantale, Program metrics} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

\funding{This work has been supported by the DIAPASoN project}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{%I want to thank \dots
}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
We study the interpretation of the lambda-calculus in a framework based on tropical mathematics, by considering the relational semantics weighted on the tropical semiring. The aim is to provide a framework where both program metrics, based on the analysis of program sensitivity via Lipschitz conditions, and resource analysis, based on higher-order program differentiation, coexist. In order to do that, we consider a graded-types lambda-calculus and the differential lambda-calculus, both languages ensuring bounded duplications.
A typical result of such line of research is that, thanks to the tropical operations, a simply typed term is interpreted as a locally Lipschitz map, which is in turn decomposed as an inf of more and more sensitive Lipschitz ones, corresponding to the interpretations of the approximants given by the lambda-calculus Taylor expansion.
We finally describe some promising lines of research, showing in particular how such tropical model expresses natural optimisation problems of probabilistic calculi.
\end{abstract}

\section{Introduction}
\input{section1}

\section{Preliminaries and motivation}
\input{section2bis}

\section{The category $\LREL$}\label{section3}

\subsection{... As tropical linear/non-linear algebra}

For $Q$ a continuous semiring, one lets the category $\QREL$ (\cite{Manzo2013} calls it $Q^\Pi$) have sets as objects and set-indexed matrices with coefficients in $Q$ as morphisms, i.e.~$\QREL(X,Y):=Q^{X\times Y}$.
%The identity morphism of $\QREL$ is the identity matrix $I\in Q^{X\times X}$,
% given by $i_{a,a}=1$ and $i_{a,b\neq a}=0$, and
The composition $st\in Q^{X\times Z}$ of $t\in Q^{X\times Y}$ and $s\in Q^{Y\times Z}$ is given by $(st)_{a,c}:=\sum_{b\in Y} s_{b,c}t_{a,b}$ (such series converges because $Q$ is continuous).
As expected, $Q^X$ is a $Q$-semimodule and we identify $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$, which have shape $f(x)_b=:\sum_{a\in X} \matr f_{a,b}x_a$, for some matrix $\matr f\in Q^{X\times Y}$.
% \begin{remark}
 %Following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we chose to see a matrix $t$ from $X$ to $Y$ as a map $t:X\times Y\to Q$.
% 
% fix $\HOM{\QREL}{X}{Y}:=Q^{X\times Y}$ with composition $st:X\times Z\to Q$ of $s:Y\times Z\to Q$ and $t:X\times Y\to Q$ defined by $(st)_{a,c}:=\sum_{b\in Y} s_{b,c}t_{a,b}$.
Notice that usual linear algebra conventions correspond to work in $\QREL^{\op}$, %a matrix $X\times Y\to Q$ is usually called a ``$Y\times X$-matrix'', meaning $Y$ rows and $X$ columns
e.g.\ the usual matrix-vector product defines a map $Q^Y\to Q^X$.
Following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we are instead working with transpose matrices.
%\end{remark}
%
%In $\QREL^{op}$ (which corresponds to systematically taking transpose matrices), composition coincides with the product matrix/matrix and $\hat{(\cdot)}$ with the product matrix/vector.
%In order to avoid confusion, we will refer to a $t\in Q^{X\times Y}$ just as a \emph{matrix from $X$ to $Y$}.
%
%
%
%We must fix a convention for matrices: following \cite{Manzo2013, Hofmann2014, Ehrhard2005}, we fix $\HOM{\QREL}{X}{Y}:=Q^{X\times Y}$ with composition $st:X\times Z\to Q$ of $s:Y\times Z\to Q$ and $t:X\times Y\to Q$ defined by $(st)_{a,c}:=\sum_{b\in Y} s_{b,c}t_{a,b}$.
%In linear algebra, a map $X\times Y\to Q$ is usually called a ``$Y\times X$-matrix'', meaning $Y$ rows and $X$ columns.
%In particular, the product of such a matrix for a vector defines a map $Q^Y\to Q^X$.
%Instead, we prefer to see a $t\in\HOM{\QREL}{X}{Y}$ as giving rise to a map $\hat t:Q^X\to Q^Y$ defined by $\hat t(x)_a:=\sum_{b\in Y} t_{a,b}x_a$.
%In $\QREL^{op}$ (which corresponds to systematically taking transpose matrices), composition coincides with the product matrix/matrix and $\hat{(\cdot)}$ with the product matrix/vector.
%In order to avoid confusion, we will refer to a $t\in Q^{X\times Y}$ just as a \emph{matrix from $X$ to $Y$}.
%As it is expected, $Q^X$ is a $Q$-semimodule and the bijection $\hat{(\cdot)}$ identifies $\HOM{\QREL}{X}{Y}$ with the set of linear maps from $Q^X$ to $Q^Y$.

The main actor of this paper is the category $\LREL$. % of sets and matrices with values over $\Lawv$ (which is a continuous semi-ring).
%, where $\Lawv$ is the already introduced Lawvere quantale, seen as the idempotent complete semiring $(\BB R_{\geq0}\cup\set{\infty},\inf,\infty,+,0)$.
%The category $\LREL$ is well-defined because $\Lawv$ is a continuous semiring (w.r.t.\ its quantale order $\preceq$.
%This amounts to check that $\min$ and $+$ commute with the $\inf$ (as operations on $\BB R_{\geq0}\cup\set{\infty}$, which is immediate), and that $(\Lawv,\preceq)$ is a cpo with $\infty$ as bottom element (which is immediate since in $\Lawv$ we have $\vee = \inf$) .
It is worth observing that the composition in $\LREL$ %is the tropicalisation of the one defining it in $\QREL$, i.e.
reads as \ $(st)_{a,c}:=\inf_{b\in Y}\set{s_{b,c}+t_{a,b}}$;
similarly, the linear functions $f:\Lawv^X\to \Lawv^Y$, %induced by matrices
which we call \emph{tropical linear}, are exactly those of shape $f(x)_b=\inf_{a\in X} \set{\matr f_{a,b}+x_a}$, for some $\matr f\in\Lawv^{X\times Y}$.
%\end{remark}

Tropical Laurent series arise from a general categorical construction:
it is well known that $\LREL$ admits a comonad $!$ which acts, on objects, by taking the finite multisets, and remember that the coKleisli category $\C C_!$ of a category $\C C$ w.r.t.\ a comonad $!$ is the category whose elements are the same of $\C C$, and $\HOM{\C C_!}{X}{Y}:=\HOM{\C C}{!X}{Y}$, with composition $\circ_!$ defined by making use of the co-multiplication of $!$.
%We will explicit this constructions in our tropical setting in the next lines.
Now, although a matrix $t\in\HOM{\LREL_!}{X}{Y}$ yields a linear map $\Lawv^{!X}\to\Lawv^Y$, by exploiting the coKleisli structure we can also ``express it in the base $X$'', i.e.\ see it as a \emph{non-linear} map $t^!:\Lawv^X\to\Lawv^Y$, by setting $t^!(x):=t\circ_! x$. %(we are identifying $\Lawv^X$ with the set $\HOM{\LREL_!}{\emptyset}{X}$ of the \emph{points} of $X$).
Concretely, we have $t^!(x)_b=\inf_{\mu\in !X} \set{\mu x+ t_{\mu,b}}$, where $\mu x:=\sum_{a\in X} \mu(a)x_a$.
%\begin{remark}%[Tropical Laurent series]
These functions corresponds then to tLs with \emph{possibly infinitely} many variables (in fact, as many as the elements of $X$), %In the following we will also refer to them as tLs. 
% We will call them simply \emph{tropical Laurent series (tLs)}.
% %Since in the general case of $\QREL$, $t^!$ is a Laurent series with operations in $Q$, let us call \emph{tropical Laurent series} the functions of shape $t^!$ for some $t\in\HOM{\LREL_!}{X}{Y}$.
%\end{remark}
%
%We find the usual notion of tLs of one variable as follows:
%\begin{remark}
and by identifying $!\set{*}\simeq \N$ and $\Lawv^{\set{*}}\simeq\Lawv$, the tLs generated by the morphisms in $\HOM{\LREL_!}{\set{*}}{\set{*}}$ are exactly the %functions $f:\Lawv\to\Lawv$ of shape $f(x)=\inf_{n\in\N}\set{nx+\matr f(n)}$, for some $\matr f:\N\to\Lawv$, i.e.\ we recover the 
usual tLs's of one variable.
% \end{remark}

In a similar way, tLs
$f:\Lawv^X\to\Lawv^Y$ 
 with \emph{finite} supports $\C F_b=\set{\mu\in!X\mid\matr f_{\mu,b}\neq\infty}$, and which have thus shape $f(x)_b=\min_{\mu\in\C F} \set{\mu x+ t_{\mu,b}}$, are generalisation of usual tropical polynomials to the case of possibly infinitely many variables.

%\end{align}
%where $+$ is the multiset union.

%Remember that in $\Lawv$ the neutral element for addition is $\infty$ and the neutral for multiplication is $0$, so for instance the evaluation map is the matrix $\RM{eval}^{X,Y}\in\Lawv^{!((X\multimap Y) \times X)\times Y}\simeq\Lawv^{(!!X\times !Y\times !X)\times Y}$ given by $\RM{eval}^{X,Y}_{\rho_1\oplus\rho_2\oplus\mu,b}:=0$ if $\rho_1=[\mu]$ and $\rho_2=[b]$, and $\RM{eval}^{X,Y}_{\rho_1\oplus\rho_2\oplus\mu,b}:=\infty$ otherwise.

\begin{remark}
 Let $X$, $Y$ be sets and let $\langle \_,\_\rangle:X\times Y \to \mathbb{R}$.
 For $f:X\to \mathbb R$, define $f^*:Y\to \mathbb R$ by $f^*(y):= \sup_{x\in X}\{\langle x,y\rangle - f(x)\}$.
 Then for $X=!A$, $Y=\Lawv^A$, where $A$ is a set, and $\langle \mu, y \rangle:= \mu y$, we have that $f^!(y)=(-f)^*(-y)$ for all $f\in\Lawv^{!A}$.
 This is precisely the same formal construction yielding the well-known convex conjugate $f^*$ of a function $f$, by taking $X$ any vector space, $Y$ its dual space, and $\langle \_,\_\rangle$ the application bilinear form (acting as the scalar product on coordinates).
 This construction is in turn a generalisation of the Legendre transformation.
 Despite the formal constructions being the same, we ignore for the moment if these could be connected to the study of high-order programs in our setting.
\end{remark}

\subsection{... As tropical weighted relational semantics for high-order languages}

The category $\LREL$ satisfies many nice properties which make it able to model all the variants of the $\lambda$-calculi to which we are interested in here. Although the majority of such properties are shared with the general class of models of $\lambda$-calculi called \emph{weighted relational models}, thoroughly studied in \cite{Manzo2013}, we mention them for our particular case in order to highlight how these look like in the tropical setting, but also to make the non-expert in such topics reader able to follow.
The real interest in our particular model will come in the next \autoref{sec:tls}.

\paragraph*{Prohibited duplication/erasure: the \emph{linear} $\STLC$}\label{sec:3A}
\input{section3A}

\paragraph*{Allowed duplication/erasure: the $\STLC$ and the probabilistic $\lambda$-calculus $\STLC_\oplus$}\label{sec:STLC+STLCO}

\input{section3B}

\input{section2}

\section{Tropical Laurent Series}\label{sec:tls}
\input{section4}

%\section{Future lines of research}\label{sec:future}

%\subsection{Quantitaive analysis}\label{sec:app}
%%\input{section5C}

%\subsection{Generalized Metric Spaces and Modules over the Lawvere Quantale}\label{section6}
%%At the basis of our approach is the observation that the \emph{tropical semiring} $([0,\infty], \min, +)$, which is at the heart of tropical mathematics, coincides with the \emph{Lawvere quantale} $\Lawv=([0,\infty], \geq, +)$ \cite{Hofmann2014, Stubbe2014}, the structure at the heart of the categorical study of metric spaces initiated by Lawvere himself \cite{Lawvere1973}.
%Let us recall that a quantale is a complete lattice endowed with a continuous monoid action. In the case of $\Lawv$ the lattice is defined by the reverse order $\geq$ on $\BB R$, and the monoid action is provided by addition. Notice that the lattice join operation of $\Lawv$ coincides with the idempotent semiring operation $\min$. 
%A consequence of these observations is that, as we discussed below, the tropical approach to linear algebra coincides with the study of ``$\Lawv$-valued matrices'', i.e.~of maps of the form $s: X\times Y\to \Lawv$ .
%In particular, a (possibly $\infty$) metric on a set $X$ is nothing but a ``$\Lawv$-valued square matrix'' $d:X\times X\to \Lawv$ satisfying axioms like e.g.~the triangular law (indeed, such distance matrices correspond to $\Lawv$-\emph{enriched categories}, a viewpoint we explicitly take in Section \ref{section6}). 
%\input{section6}%

\section{Related Work}\label{section7} \input{section7}

\section{Conclusion and future work}\label{section8} \input{section8}

%%
%% Bibliography
%%

%% Please use bibtex,

%\bibliography{tropical}

%\appendix

%\section{\autoref{section3}}\label{app-sec:section3}

\end{document}