### 2023-07-28 CSL24/section2bis.tex

\subsection{Elements of tropical mathematics}
\input{section3new}

%\subsection{Tropical non-linear algebra}\label{sec:4A}

\begin{definition}
A \emph{tropical power series} (tPs) of one variable is a function $f:\Lawv\to\Lawv$ of shape $f(x)=\inf_{j\in I}\{i_j x+c_{i_{j}}\}$, with $i_{j}\in\N$, $c_{i_{j}}\in\Lawv$ and $I\subseteq \N$.
If $I$ is finite, $f$ is said to be a \emph{tropical polynomial}. 
\end{definition}

A tropical polynomial is always a piece-wise linear function since it has shape $f(x)=\min_{0\leq j\leq n}\{i_{j}x+c_{i_{j}}\}$
For example, the polynomials $\varphi_{n}(x)=\min_{0\leq j\leq n}\{jx+2^{-j}\}$
are illustrated in Fig.~\ref{fig:plot1} for $0\leq n \leq 4$.

%\begin{figure}
%\begin{subfigure}{0.4\textwidth}
%\begin{tikzpicture}[scale=0.6]
%\begin{axis}[samples=250]
%\addplot[yellow,domain=0:0.8] {1+0.02};
%
%\addplot[orange,domain=0:0.8] {min(x+1/2, 1)+0.01};
%
%
%\addplot[red,domain=0:0.8] {min(2*x+1/4, x+1/2, 1};
%\addplot[blue,domain=0:0.8] {min(3*x+1/8,2*x+1/4, x+1/2, 1)-0.01};
%\addplot[orange,domain=0:0.8] {min(4*x+1/16,3*x+1/8,2*x+1/4, x+1/2, 1)-0.02};
%
%\addplot[violet,domain=0:0.8] {min(
%10*x+1/1424,
%9*x+1/712,
%8*x+1/356,
%7*x+1/128,
%6*x+1/64,
%5*x+1/32,
%4*x+1/16,3*x+1/8,2*x+1/4, x+1/2, 1)-.03};
%
%
%\end{axis}
%
%\end{tikzpicture}
%\caption{}
%\label{fig:plot1}
%\end{subfigure}
%\begin{subfigure}{0.4\textwidth}
%\begin{tikzpicture}[scale=0.6]
%\begin{axis}[samples=50, view={15}{45}]
%\addplot3[orange,domain=0:2] {min(2*x, 2*x+y, 3*y)};
%
%\end{axis}
%\end{tikzpicture}
%\caption{}
%\label{fig:plot2}
%\end{subfigure}
%\label{fig:plot12}
%\caption{
%\ref{fig:plot1}
% (a) Plot of the tropical polynomials $\varphi_{n}$, for $0\leq n\leq 4$ (from top to bottom), and of their limit tLs $\varphi$ (in violet). The points where the slope changes are  the tropical roots of $\varphi$, i.e.~the points $x=2^{-(i+1)}$, satisfying $ix+2^{-i}=(i+1)x+2^{-(i+1)}$.
%\ref{fig:plot2}
%(b) Plot of $\varphi(x,y)=\min\{2x, 2x+y,3y\}$.
%}
%\end{figure}

\begin{wrapfigure}{r}{0.5\textwidth}%\begin{figure}
\begin{tikzpicture}[scale=0.9]
\begin{axis}[samples=250]
\addplot[yellow,domain=0:0.8] {1+0.02};

\addplot[orange,domain=0:0.8] {min(x+1/2, 1)+0.01};

\addplot[red,domain=0:0.8] {min(2*x+1/4, x+1/2, 1};
\addplot[blue,domain=0:0.8] {min(3*x+1/8,2*x+1/4, x+1/2, 1)-0.01};
\addplot[orange,domain=0:0.8] {min(4*x+1/16,3*x+1/8,2*x+1/4, x+1/2, 1)-0.02};

\addplot[violet,domain=0:0.8] {min(
10*x+1/1424,
9*x+1/712,
8*x+1/356,
7*x+1/128,
6*x+1/64,
5*x+1/32,
4*x+1/16,3*x+1/8,2*x+1/4, x+1/2, 1)-.03};

\end{axis}

\end{tikzpicture}
\caption{\small Tropical polynomials $\varphi_0,\dots,\varphi_4$ (top to bottom), and the limit tLs $\varphi$ (in violet). The points where the slope changes are  the tropical roots of $\varphi$, i.e.~the points $x=2^{-(i+1)}$, satisfying $ix+2^{-i}=(i+1)x+2^{-(i+1)}$.}
\label{fig:plot1}%\end{figure}%
\end{wrapfigure} %%

A \emph{tropical root} of a tropical power series $\varphi$ is a point $x\in \Lawv$ where $\varphi$ is not differentiable (i.e.~where the slope of $\varphi$ changes), or equivalently, where the minimum defining $\varphi$ is attained at least twice (see \autoref{fig:plot1}).
%For instance, the tropical roots of $\varphi_{n+1}$ are of the form $2^{-(i+1)}$, for $0\leq i \leq n$.
%Tropical roots yield the usual factorization property of roots: if $x_{0}$ is a root of $f$, this factorizes as $f(x)=k\min\{x,x_{0}\}+ g(x)$.
Unlike in standard algebra, tropical roots can be computed in linear time \cite{Noferini2015}.
%Tropical polynomials and their roots are the main object of study of tropical geometry.
%
%A
% simple calculation shows that this condition corresponds, in the tropical setting, to the usual notion of root.

Since $\inf$s are not in general $\min$s, a tPs is a ``limit'' of tropical polynomials of higher and higher degree, and its behaviour may be less predictable than that of tropical polynomials~\cite{Porzio2021}. %For instance, tropical roots for tLs (see \cite{Porzio2021}) may also include limit points of their domain.
E.g., the tLs $\varphi(x):=\inf_{n\in\N}\set{nx+2^{-n}}$ (see Fig.~\ref{fig:plot1}) %that we will take as our running example, 
is the ``limit'' of the polynomials $\varphi_{n}$.

%
%
%We will take as our running example
%%given respectively by Equation~\ref{eq:polytrop} and Equation~\ref{eq:defTLS}.
%A positive $x\in[0,\infty)$ is said to be a (finite) \emph{tropical root} of a tropical Laurent series $f:\Lawv\to\Lawv$ iff $f$ is not differentiable at $x$ (w.r.t.\ the usual topology on $\BB R_{\geq0}$.This is equivalent to ask that the $\inf$ defining $f$ at $x$ is obtained \emph{at least} twice.{\color{red} \`e vero anche per TLS o solo per poly questo? Inoltre, Robol parla anche di finite end-points of the domain: nel nostro caso sarebbe $x=0$}

%\begin{example}\label{ex:famous_ex}
% The function $f:\Lawv\to\Lawv$ defined by $f(x):=\inf\limits_{n\in\N}\set{nx+\frac{1}{2^n}}$ is a tropical Laurent series.
%\end{example}

%By plotting its graph {\color{red}vogliamo plottarlo?}, we observe several properties that we will lift to the more general case that we will consider in the next lines, and this example will serve as running one.

\begin{remark}\label{rmk:val_trop}
 There is a clear and well-known relation between tropical polynomial/power series and usual polynomials/power series. 
Given $Q,L$ semirings with units and zeros, $L$ commutative idempotent, if $L$ is totally ordered by the partial order: $\alpha \preceq \beta$ iff $\alpha +  \beta = \beta$, then one defines a \emph{valuation} \cite{Izhakian2015} to be a map $\mathrm{val}:Q\to L$ s.t.\ $\mathrm{val}(0)=0$, $\mathrm{val}(1)=1$, $\mathrm{val}(ab)=\mathrm{val}(a)\mathrm{val}(b)$, $\mathrm{val}(a+b)\preceq \max\set{\mathrm{val}(a),\mathrm{val}(b)}$.
The valuation which is constantly equal to $1\in L$ on all the invertibles of $Q$ is called the \emph{trivial valuation} $\mathrm{val}_1$.

Now, for a (formal) power series $f=\sum_{i\in I} a_i x^i$ (polynomials being the case for $I$ finite) with coefficitents $a_i$ in a semiring $Q$, and for a fixed valuation $\mathrm{val}:Q\to \Lawv$ (notice that in $\Lawv$ we have that $\preceq$ is $\geq$ and $\max^{\preceq}=\min$), one defines the \emph{tropicalisation} $\trop^{\mathrm{val}}f:\Lawv \to \Lawv$ of $f$ as the tropical polynomial/power series function $\trop^{\mathrm{val}}f(\alpha):=\inf_{i\in I} \set{\mathrm{val}(a_i)+i\alpha}$.

Remark that tropicalising w.r.t.\ the trivial valuation gives $\trop^{\mathrm{val}_1} \sum_{i\in I} a_i x^i = \inf_{i\in I} \set{i\alpha} = \trop^{\mathrm{val}} \sum_{i\in I\subseteq\N} x^i$ for all valuation $\mathrm{val}$, because in $\Lawv$ the unit is $0$. %and the zero is $\infty$.
%Remember that a_i \neq 0 by def of formal power series
That is, it coincides to consider the same power series but with coefficients in $Q$ quotiented for the idempotency of $+$.
\end{remark}

\subsection{Log-Probabilities and Tropicalisation through an exemple}\label{sec:proba}
\input{section5}