### 2023-07-28 CSL24/section3A.tex

$\LREL$ is a Symmetric Monoidal Closed Category (SMCC) \cite[Section III.A]{Manzo2013}, %(actually, even a model of IMLL), 
whose monoidal product $\otimes$ acts on objects as the Cartesian product of sets, %and as the \emph{Kronecker product} of matrices on morphisms.
 %{(\color{red} Ref?)} 
 and its internal hom $X\multimap Y$ is also given by the Cartesian product $X\times Y$ of sets, with the usual evaluation and curry maps.
As such, it is able to interpret the \emph{linear simply typed $\lam$-calculus}, the restriction of the ordinary $\lam$-calculus in which programs ,% following the linear logic mantra, 
can only use their arguments \emph{exactly once}. %(and indeed this calculus can be embedded in its \emph{intuitionistic multiplicative} fragment \emph{IMLL}).
%More precisely, the linear $\lam$-calculus is obtained from the ordinary one by adding the constraint that each $\lam$-abstracted variable appears exactly once in the scope of the $\lam$-abstraction.
%In order to frame it in a category $\C C$, one needs a symmetric tensor product $\otimes$ together with internal hom-set objects $X\multimap Y$ s.t.\ the evaluation and curry maps yield a \emph{symmetric monoidal adjunction}: $\C C(Z\otimes X, Y) \simeq \C C (Z, X\multimap Y)$.