### 2023-07-29 CSL24/sectionLaw.tex

As we have seen, the morphisms of $\LREL$ can be seen as continuous functions between the $\Lawv$-modules $\Lawv^{X}$, when the latter are taken with the metric induced by the $\infty$-norm. This viewpoint gives a metric flavor to $\LREL$, and allowed us to relate differential and metric structure. Yet, how far can this correspondence be pushed?
In particular, is this correspondence restricted to $\Lawv$-modules of the form $\Lawv^{X}$ (i.e.~with a fixed base), or does it hold in some sense for arbitrary $\Lawv$-modules? Is this correspondence restricted to the $\infty$-norm metric, or does it hold for other metrics too?

An answer to these questions comes from an elegant categorical correspondence between tropical linear algebra and the theory of \emph{generalized metric spaces}, initiated by Lawvere's pioneering work \cite{Lawvere1973}, and at the heart of the emergent field of \emph{monoidal topology} \cite{Hofmann2014, Stubbe2014}, that it's worth recalling.

%In this section we first reconstruct this correspondence, by combining some {folklore} results with more abstract ones from recent literature in enriched category theory \cite{Fuji, Stubbe2006, Shen2014}. Then we show that this correspondence can be lifted to a model of the full differential $\lambda$-calculus, by suitably generalizing the construction of the Lafont exponential of $\LREL$.
%
%
%also extends exploit this correspondence to prove that arbitrary $\Lawv$-modules (equivalently, arbitrary \emph{complete} generalized metric spaces) provide a model of the differential $\lambda$-calculus, hence generalizing the tropical relational model.
%

%{$\Lawv$-modules}

On the one hand we have $\Lawv$-modules: these are triples $(M,\preceq, \star)$ where $(M, \preceq)$ is a sup-lattice, and $\star: \Lawv \times M \to M$ is a continuous (left-)action of $\Lawv$ on it, where continuous means that $\star$ commutes with both joins in $\Lawv$ and in $M$.% (notice that this is indeed the usual notion of module over the Lawvere quantale $\Lawv$).
A $\Lawv$-module homomorphism is a map $f:M\to N$ commuting with both joins and the $\Lawv$-action. We let $\Mod$ indicate the category of $\Lawv$-modules and their homomorphisms.

% 
% 
%
%$\Lawv$ is the most basic example of $\Lawv$-module.
%Any $\Lawv$-module $M$ has a dual $M^{\op}$, with reversed order and (right-)action $x\multimapinv \epsilon= \bigvee\{y\mid \epsilon \star y\succeq x\}$.
% Other basic examples of $\Lawv$-modules are the sets $\Lawv^{X}$, with order and action defined pointwise. 
% 
% 
% While the $\Lawv^{X}$ have a fixed base, for an arbitrary $\Lawv$-module one can retrieve a base via the \emph{Yoneda embedding}
%$\Yon: M \to \Mod(M^{\op},\Lawv)$, where $\Yon(x)(y)=\inf\{\epsilon\mid \epsilon\star y\succeq x\}$. 
%
%
%\begin{proposition}[cf.~\cite{Stubbe2014, Shen2014}]\label{prop:yonemod}
%For any $\Lawv$-module $M$, the Yoneda embedding has a left-adjoint $\sup(f)=\bigvee_{x\in M}f(x)\star x$.
%%\end{proposition}
%
%
%Like $\LREL$, the category $\Mod$ has the relevant structure to interpret the linear $\lambda$-calculus:
%\begin{proposition}
%$\Mod$ is a SMCC.
%\end{proposition}
%%\begin{proof}
%The hom-sets $\Mod(M,N)$ have a natural $\Lawv$-module structure, defined pointwise. The tensor product of $\Lawv$-modules $M\otimes N$ can be defined as the quotient of the usual tensor product of sup-lattices under the smallest congruence containing all pairs $(\{(\epsilon \star x,y)\},\{(x,\epsilon\star y)\})$ (see e.g.~\cite{Russo2007}).
%Notably, any element of $M\otimes N$ can be identified with a join of basic tensors $x\otimes y$, corresponding to the equivalence class of the pair $\langle x,y\rangle$.
%Beyond the required adjointness of the internal hom and the tensor, one can check that $\Mod$ is actually \emph{$^{*}$-autonomous}, since it satisfies $(M^{\op})^{\op}\simeq M$ and 
%$\Hom(M,N^{\op})\simeq (M\otimes N)^{\op}$.
%Finally, $\Mod$ has \emph{biproducts}: products and coproducts are both given by the Cartesian product of the underlying posets, with action defined pointwise.
%
%, similarly to the case of sup-lattices \cite{}, as the quotient of the free sup-lattice $\C P(M\times N)$ under the smallest congruence containing all pairs $((\bigvee A,y),\bigcup_{a\in A}\{(a,y)\})$, for $A\subseteq M, y\in N$, 
%$((x,\bigvee B),\bigcup_{b\in B}\{(x,b)\})$, for all $x\in A$, $B\subseteq N$, and all pairws $\{
%\end{proof}
%
%
%\begin{remark}
%The SMCC structure of $\Mod$ coincides with that of $\LREL$ for the modules $\Lawv^{X}$: we already know that $\Mod(\Lawv^{X},\Lawv^{Y})\simeq \Lawv^{X\times Y} $, and 
%one can prove $\Lawv^{X}\otimes \Lawv^{Y}\simeq \Lawv^{X\times Y} $ (cf.~\cite{Russo2007}).% and 
%%$\Pi_{i\in I}\Lawv^{X_{i}}\simeq \Lawv^{\coprod_{i\in I}X_{i}}$.
%\end{remark}

%\subsection{$\Lawv$-categories}
On the other hand we have Lawvere's \emph{generalized metric spaces}:
Lawvere was the first to observe that a metric space can be described as a \emph{$\Lawv$-enriched} category. Indeed, spelling out the definition, a $\Lawv$-enriched category (in short, a $\Lawv$-category) is given by a set $X$ together with a ``hom-set'' $X(-,-):X\times X\to \Lawv$ satisfying 
$0  \geq X(x,x)$ and $X(y,z)+X(x,y)&\geq  X(x,z)$. This structure is often referred to as a \emph{generalized metric space} \cite{Lawvere1973, Hofmann2014, Stubbe2014}. %, or as a \emph{quasi-metric space}. 
 A metric space, in the usual sense, is thus the same as a skeletal (i.e.~$X(x,y)=0$ implies $x=y$) and symmetric (i.e.~$X(x,y)=X(y,x)$) $\Lawv$-category.
Notice that a $\Lawv$-enriched \emph{functor} between $\Lawv$-categories is just a non-expansive map $f:X\to Y$, as functoriality reads as $Y(f(x),f(y))\leq X(x,y)$.

A $\Lawv$-category is \emph{complete} when all weighted colimits (cf.~Appendix for a precise definition) exist in $X$. This categorical notion of completeness can be seen as a generalization of the usual notion of Cauchy-completeness.
We let $\GMet$ indicate the category of complete $\Lawv$-categories and continuous (i.e.~colimit-preserving) $\Lawv$-enriched functors.

Now, it turns out that the categories $\Mod$ and $\GMet$ are isomorphic \cite{Stubbe2006}. First,  any $\Lawv$-module $(M,\preceq, \star)$ is a $\Lawv$-category via
$M(x,y) = \inf\left\{ \epsilon \mid \epsilon \star x\geq y\right\}$. Moreover, a homomorphism of $\Lawv$-modules is a functor of the associated $\Lawv$-categories. Conversely, complete $\Lawv$-categories are \emph{tensored}, that is, admit a $\Lawv$-module structure $\epsilon \otimes x$ (see the Appendix).
Moreover, a continuous functor between complete $\Lawv$-categories is the same as a homomorphism of the associated $\Lawv$-modules. 
%
%
%To conclude our correspondence between $\Lawv$-modules and complete $\Lawv$-categories, it remains to observe that the 
%two constructions leading from one structure to the other are one the inverse of the other: for any $\Lawv$-module $(M,\preceq,\star)$,
%$x\preceq_{M}y$ iff $M(y,x)=0$ iff $x\preceq y=0\star y$, and, from  
%$M(\epsilon \star x, y)= M(x,y)\dotdiv \epsilon$, we deduce $\epsilon\otimes x=\epsilon \star x$. 
%Conversely, 
%for any complete $\Lawv$-category $X$ and $x,y\in X$, one can check that 
%$X(y,x)=\inf\{ \epsilon \mid X(\epsilon\otimes y,x )=0\}$.
%
%This leads to the following:
%
%
%\begin{theorem}[cf.~\cite{Stubbe2006}]
%$\Mod$ and $\GMet$ are isomorphic categories.
%\end{theorem}

%
%Since $\Mod$ (and thus $\GMet$ too) is a SMCC, it is worth making its metric structure explicit. Given complete $\Lawv$-categories $X,Y$, we have that the distance on the hom-set $\Mod(X,Y)$ is given by 
%\begin{align}
%\Mod(X,Y)(f,g)= \sup_{x\in X}Y(f(x),g(x));
%\end{align}
%and the distance on $X\otimes Y$ is given by
%\begin{align}
%(X\otimes Y)(\alpha, \beta)=
%\sup_{i}\inf_{j}\left\{X(x_{i},x'_{j})+Y(y_{j},y'_{j}\right\},
%\end{align}
% where $\alpha=\bigvee_{i}x_{i}\otimes y_{i}$ and 
%$\beta=\bigvee_{j}x'_{j}\otimes y'_{j}$, 
%from which we deduce that, for basic tensors
%$x\otimes y, x'\otimes y'$, their distance is just $X(x,x')+Y(y,y')$.

%
%
%and is extended to arbitrary joins by continuity, i.e.~
%$(X\otimes Y)(\alpha, \beta)=
%\sup_{i}\inf_{j}X(x_{i},x'_{j})+Y(y_{j},y'_{j}),
%$, where $\alpha=\bigvee_{i}x_{i}\otimes y_{i}$ and 
%$\beta=\bigvee_{j}x'_{j}\otimes y'_{j}$.
%%and thus coincides with the sum-metric over basic tensors;
%\item the distance on the bi-product $\Pi_{i\in i}X_{i}$ is given by
%\begin{align}
%(\Pi_{i\in I}X_{i})(f,g)= \sup_{i\in I}X_{i}(f(x),g(x));
%\end{align}

%\end{itemize}

To conclude, we show how the correspondence $\Mod\simeq \GMet$ lifts to a model of the differential $\lambda$-calculus, extending the co-Kleisli category $\LREL_{!}$.

First, we need to define a Lafont exponential $!$ over $\Mod$, and for this we use a well-known recipe from \cite{Mellies2018, Manzo2013}, that is: we first define a symmetric algebra $\Sym_{n}(M)$ as the equalizer of all permutative actions on $n$-tensors $M\otimes \dots \otimes M$; then, exploiting suitable properties of $\Mod$, we may define $!$ as the product of the symmetric algebras.  
This ensures that the coKleisli category $\Mod_{!}$ is a CCC.
Moreover, the constructions for $\Mod$ generalize those of $\LREL$, in the sense that $!(\Lawv^{X})\simeq \Lawv^{\multiset(X)}$ and that $\Mod_{!}(\Lawv^{X},\Lawv^{Y})\simeq \LREL_{!}(X,Y)$.

Finally, general results {\color{red}CITARE} allow us to conclude 
\begin{theorem}
$\Mod_{!}\simeq\GMet_{!}$ is a CC$\partial$C, when equipped with 
the differential operator $E$ defined, for $f:!M\to N$, by 
$Ef(\alpha)=
\bigvee\left\{
f(\beta\cup [x]) \ \Big \vert  \ 
\iota_{n}(\beta)\otimes \iota_{1}(x) \leq S(\alpha)
\right\}
$, where $\iota_{k}: M_{k}\to \prod_{i\in I}M_{i}$ is the injection morphism given by $\iota_{k}(x)( k)=x$ and $\iota_{k}(x)(i\neq k)=\infty$,
and $S: !(M\times N)\to !M\otimes !N$ is the Seely isomorphism \cite{Mellies2018}.
\end{theorem}