### 2023-11-27 CSL24/ArXiv/section7.tex

The applications of tropical mathematics in computer science abound, e.g.~in automata theory \cite{Chua1992, Simon}, machine learning \cite{Maragos2021, Pachter2004, Zhang2018}, optimization \cite{Akian2011, Akian2012}, and convex analysis \cite{Lucet2009}.

As we said, the relational semantics over the tropical semiring was quickly explored in \cite{Manzo2013}, to provide a ``best case'' resource analysis of a $\mathrm{PCF}$-like language with non-deterministic choice. 
The connections between differential $\lambda$-calculus (and differential linear logic), relational semantics, and non-idempotent intersection types are very well-studied (see \cite{decarvalho2018}, and more recently, \cite{Mazza2016} for a more abstract perspective, and \cite{Olimpieri2021, Galal2021} for a 2-categorical, or proof-relevant, extension).
\emph{Probabilistic coherent spaces} \cite{Ehrhard2011}, a variant of  the relational semantics, provide an interpretation of higher-order probabilistic programs
as analytic functions. In \cite{Ehrhard2022} it was observed that such functions satisfy a local Lipschitz condition somehow reminiscent of our examples in Section \ref{section2}.

The study of linear or bounded type systems for sensitivity analysis was initiated in \cite{Girard92tcs} and later developed \cite{Schopp, SchoppDalLago, Reed2010}.
%As recalled in the paper, the use bounded exponentials ensures that well-typed programs satisfy a Lipschitz condition.
Related approaches, although not based on metrics, are provided by \emph{differential logical relations} \cite{dallago} and \emph{change action} models \cite{Picallo2019}.
More generally, the literature on program metrics in denotational semantics is vast. Since at least \cite{VANBREUGEL20011} metric spaces, also in Lawvere's generalized sense \cite{Lawvere1973}, have been exploited as an alternative framework to standard, domain-theoretic, denotational semantics; notably models of $\STLC$ and $\mathrm{PCF}$ based on 
\emph{ultra}-metrics and \emph{partial} metrics have been proposed  \cite{Escardo1999,PistoneLICS, PistoneFSCD2022}.

Motivated by connections with computer science and fuzzy set-theory, 
the abstract study of generalized metric spaces in the framework of \emph{quantale}- or even \emph{quantaloid}-enriched categories has led to a significant literature in recent years (e.g.~\cite{Hofmann2014, Stubbe2014}), 
and connections with tropical mathematics also have been explored e.g.~in \cite{Fuji, Willerton2013}. Moreover, applications of quantale-modules to both logic and computer science have also been studied \cite{Abramsky1993b, Russo2007}.

Finally, connections between program metrics and the differential $\lambda$-calculus have been already suggested in \cite{PistoneLICS}; moreover, \emph{cartesian difference categories} \cite{Picallo2020} have been proposed as a way to relate derivatives in differential categories with those found in change action models.

%While our discussion in Section \ref{section5} is inspired by a well-known application of tropical polynomials to Hidden Markov Models \cite{Pachter2004},  
%the vast literature in this domain lets us think that other ways to 
%apply tropical semantics to the analysis of higher-order programs might be studied.
%
%
%Other connections tropical/metric -> Fuji, ??
%Applications of tropical to computer science.
%Log-probabilities.
%Quantale-modules -> Abramski?
%
%