### 2023-11-30 CSL24/main.tex
commit

\documentclass[a4paper,english, UKenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{bussproofs}
\EnableBpAbbreviations
\usepackage{hyperref}
\usepackage{tikz,pgfplots}
\pgfplotsset{compat=newest}
\usepackage{tikz-cd}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usetikzlibrary{cd}
\usepackage{mathabx}
\usepackage{cmll}
\usepackage{stmaryrd}
\usepackage{wrapfig}
\usepackage{relsize}

\usetikzlibrary{intersections,through}

\input{prooftree}

\usepackage{proof}

\bibliographystyle{plainurl}% the mandatory bibstyle

\input{macros}

\addto\extrasenglish{
  \def\sectionautorefname{Section}
}

\title{Tropical Mathematics and the Lambda-calculus I}
\subtitle{Metric and Differential Analysis of Effectful Programs
} %TODO Please add

\titlerunning{Tropical mathematics for the Lambda-calculus I} %TODO optional, please use if title is longer than one line

\author{Davide {Barbarossa}}{Dipartimento di Informatica - Scienza e Ingegneria, Universit\`a di Bologna, %[optional: Address], 
Italy% \and My second affiliation, Country 
\and \url{https://lipn.univ-paris13.fr/\~barbarossa/index.html} }{davide.barbarossa@unibo.it}{https://orcid.org/0000-0003-4608-8282}{%(Optional)Funding each author
}%TODO mandatory, please use full name; only 1 author per \author macro; first two parameters are mandatory, other parameters can be empty. Please provide at least the name of the affiliation and the country. The full address is optional. Use additional curly braces to indicate the correct name splitting when the last name consists of multiple name parts.

\author{Paolo {Pistone}}{Dipartimento di Informatica - Scienza e Ingegneria, Universit\`a di Bologna, %[optional: Address], 
Italy% \and LIP, Universit\'e de Lyon, France 
\and \url{http://logica.uniroma3.it/pistone/
} }{paolo.pistone2@unibo.it}{https://orcid.org/0000-0003-4250-9051}{%(Optional)Funding each author
}

\authorrunning{D.\,Barbarossa and P.\,Pistone} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

\Copyright{Davide Barbarossa and Paolo Pistone} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[500]{Theory of computation~Lambda calculus}
\ccsdesc[300]{Theory of computation~Categorical semantics}
\ccsdesc[100]{Theory of computation~Linear logic}%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Relational semantics, Differential lambda-calculus, Tropical semiring,  Program metrics, Lawvere quantale} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{\url{https://arxiv.org/abs/2311.15704}} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

\funding{This work has been supported by the European Research Council through the project ERC CoG 818616 DIAPASoN.}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

\acknowledgements{%I want to thank \dots
}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{Aniello Murano and Alexandra Silva} 
\EventNoEds{2} 
\EventLongTitle{32nd EACSL Annual Conference on Computer Science Logic (CSL 
2024)} 
\EventShortTitle{CSL 2024} 
\EventAcronym{CSL} 
\EventYear{2024} 
\EventDate{February 19--23, 2024} 
\EventLocation{Naples, Italy} 
\EventLogo{} 
\SeriesVolume{288} 
\ArticleNo{35}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

%TODO mandatory: add short abstract of the document
\begin{abstract}
We study the interpretation of the lambda-calculus in a framework based on tropical mathematics, and we show that it provides a unifying framework for two well-developed quantitative approaches to program semantics: on the one hand program metrics, based on the analysis of program sensitivity via Lipschitz conditions, on the other hand resource analysis, based on linear logic and higher-order program differentiation. 
To do that, we focus on the semantics arising from the relational model weighted over the tropical semiring, and we discuss its application to the study of “best case” program behavior for languages with probabilistic and non-deterministic effects. Finally, we show that a general foundation for this approach is provided by an abstract correspondence between tropical algebra and Lawvere’s theory of generalized metric spaces.
%A typical result of such line of research is that, thanks to the tropical operations, a simply typed term is interpreted as a locally Lipschitz map, which is in turn decomposed as an inf of more and more sensitive Lipschitz ones, corresponding to the interpretations of the approximants given by the lambda-calculus Taylor expansion.
%We finally describe some promising lines of research, showing in particular how such tropical model expresses natural optimisation problems of probabilistic calculi.
\end{abstract}

\section{Introduction}
\input{section1}

\section{A Bridge between Metric and Differential Aspects}\label{section5bis}
\input{section5bis}

\section{Tropical Polynomials and Power Series}\label{section2}
\input{section2bis}

\section{Tropical Semantics and First Order Effectful Programs}\label{section22}
\input{section2bisbis}

\section{Tropical Semantics of Higher-Order Programs}\label{section3}
\input{section3fancy}

%%
%%\section{Topical linear/non-linear algebra}\label{section3}
%%\input{section3C}
%%
%%\section{Tropical weighted relational semantics for high-order languages}
%%
%%The category $\LREL$ satisfies many nice properties which make it able to model all the variants of the $\lambda$-calculi to which we are interested in here. Although the majority of such properties are shared with the general class of models of $\lambda$-calculi called \emph{weighted relational models}, thoroughly studied in \cite{Manzo2013}, we mention them for our particular case in order to highlight how these look like in the tropical setting, but also to make the non-expert in such topics reader able to follow.
%%The real interest in our particular model will come in the next \autoref{sec:tls}.
%%
%%
%%\input{section3A}
%%
%%\input{section2}
%
%\paragraph*{Allowed duplication/erasure: the probabilistic $\lambda$-calculus $\STLC_\oplus$}\label{sec:STLCO}
%
%\input{section3B}

\section{On Tropical Power Series}\label{sec:tls}
\input{section4}

\section{Lipschitz Meets Taylor}\label{sec:TayLip}
\input{sectionTaylip}

\section{Generalized Metric Spaces and $\Lawv$-Modules}\label{sec:GMS}
\input{sectionLaw}

%\section{Future lines of research}\label{sec:future}

%\subsection{Quantitaive analysis}\label{sec:app}
%%\input{section5C}

%\subsection{Generalized Metric Spaces and Modules over the Lawvere Quantale}\label{section6}
%%At the basis of our approach is the observation that the \emph{tropical semiring} $([0,\infty], \min, +)$, which is at the heart of tropical mathematics, coincides with the \emph{Lawvere quantale} $\Lawv=([0,\infty], \geq, +)$ \cite{Hofmann2014, Stubbe2014}, the structure at the heart of the categorical study of metric spaces initiated by Lawvere himself \cite{Lawvere1973}.
%Let us recall that a quantale is a complete lattice endowed with a continuous monoid action. In the case of $\Lawv$ the lattice is defined by the reverse order $\geq$ on $\BB R$, and the monoid action is provided by addition. Notice that the lattice join operation of $\Lawv$ coincides with the idempotent semiring operation $\min$. 
%A consequence of these observations is that, as we discussed below, the tropical approach to linear algebra coincides with the study of ``$\Lawv$-valued matrices'', i.e.~of maps of the form $s: X\times Y\to \Lawv$ .
%In particular, a (possibly $\infty$) metric on a set $X$ is nothing but a ``$\Lawv$-valued square matrix'' $d:X\times X\to \Lawv$ satisfying axioms like e.g.~the triangular law (indeed, such distance matrices correspond to $\Lawv$-\emph{enriched categories}, a viewpoint we explicitly take in Section \ref{section6}). 
%\input{section6}%

\section{Related Work}\label{section7} \input{section7}

\section{Conclusion and Future Work}\label{section8} \input{section8}

%%
%% Bibliography
%%

%% Please use bibtex,

\bibliography{tropical}

\appendix

\section{Proofs from \autoref{section2}: Theorem \ref{theorem:fepsilon}}
\input{appSection3}

\section{Proofs from \autoref{section3}: $(\LREL, !_{n})$ is a model of $\BSTLC$}
\input{appSection5}

\section{Proofs from \autoref{sec:tls}: the Analysis of Tropical Power Series}\label{appsec:tls}
\input{app_section4}

\section{Proofs from \autoref{sec:TayLip}: Lipschitz meets Taylor}
\input{appSection7}

\section{Proofs of \autoref{sec:GMS}: $\Lawv$-modules and Generalized Metric Spaces}
\input{appSection8}

\end{document}